% !TeX root = closed-systems-from-comparison-completeness.tex
% ============================================================
\chapter*{Introduction}
\phantomsection
\addcontentsline{toc}{chapter}{Introduction}
\label{ch:introduction}
% ============================================================
This introduction begins the formal argument.
The problem is to describe a world in which the agents, clocks,
comparison protocols, reports, and measuring conventions are themselves
internal to the object being modeled.
In such a setting, a background manifold, state-space linearity, causal
order, field content, observer frame, or measurement scale cannot be used
as the primitive source of distinctions.
Those structures may appear later, but only after the text has fixed
their status as reconstructed structure, realization data, or boundary
data.

The primitive object is deliberately small.
It is not a manifold, a Hilbert space, a field bundle, or a causal order.
It is a set of states equipped with binary comparison predicates.
The first task is to specify which predicates count as intrinsic.
The second is to determine what quotient semantics, transport residue,
smooth realization, and observer-side readings are forced or permitted
once the system is required to close over its own comparison data.

The resulting problem is a mathematical classification problem.
Which distinctions survive descent through the intrinsic symmetries?
Which structures are forced by comparison completeness and transport
obstruction?
Which familiar smooth or field-theoretic objects enter only at a
realization interface?
External-time and external-reference issues familiar from canonical
quantum gravity motivate the closure requirement
\cite{Isham1993,Kuchar1992,Anderson2017}, but the present argument uses
that literature only as context.  The formal object of study is the
comparison system defined below.

\medskip
\noindent
\emph{Guiding question.}
\emph{Given only intrinsic comparison data on a closed world, which
mathematical structures are necessary consequences rather than
auxiliary choices?}
\medskip

\section{Scope of the closure claim}
The closure claim made here is not that every mathematical structure used
in physics must be eliminable.
It is a claim about logical status.
Structures that are internally reconstructed have theorem-level status
inside the closed stack.
Structures that are needed to compare the closed stack with a smooth,
macroscopic, or observer-side description are interface data unless they
are separately derived.
The distinction is essential throughout the text.

Thus a smooth manifold may appear as a realized branch, but it is not a
primitive arena.
A scalar may be classified intrinsically, but its identification with a
particular observer-side measured mass requires a semantic readout map.
The quadratic transport carrier may be forced by the obstruction stack,
while faithful smooth realization of that carrier is analyzed through
condition \textup{(T)}, axiom \textup{(D)}, and the canonical realization
constructed in \cref{ch:interface}.
The state-side observable algebra may be larger than any realized smooth
reading; the regular sector is then selected by a stated scale-coherence
condition in \cref{ch:dynamics-program}.

The monograph therefore uses three recurring statuses.
First, \emph{intrinsic consequences} are results proved from comparison
data and closed-world admissibility.
Second, \emph{realization-level statements} are results proved after a
specified smooth, Hilbert, phase, or observer-side reading has been
introduced.
Third, \emph{boundary data} are items deliberately not absorbed into the
intrinsic theorem stack.
The conditional-completion ledger in \cref{app:conditional-completion}
records the remaining global boundary data for the numerical and
realization layers.

\section{Primitive object and admissibility stance}
\begin{definition}[Comparison world]
	\label{def:intro-comparison-world}
	A \emph{comparison world} is a pair
	\[
		(U,\mathcal C),
	\]
	where \(U\neq\varnothing\) is a set of states and
	\(\mathcal C\) is a family of binary comparison predicates
	\[
		c:U\times U\to\{0,1\}.
	\]
	We require \(\mathcal C\) to be \emph{intrinsic}: writing
	\(G=\Aut(U,\mathcal C)\) for the automorphism group of
	\cref{def:intrinsic-symmetry-group}, every predicate is
	\(G\)-invariant,
	\[
		c(gu,gv)=c(u,v)
		\qquad
		\text{for all }c\in\mathcal C,\ g\in G,\ u,v\in U.
	\]
	This is a fixed-point condition: \(G\) is determined by \(\mathcal C\),
	and \(\mathcal C\) is admissible exactly when its predicates are stable
	under the symmetries they generate. A predicate violating it assigns
	comparison values tracking a labeling of states that no symmetry of the
	world preserves; such a predicate is an external evaluator in the sense
	of \textup{(SP1)}, not an intrinsic comparison. Intrinsicality is
	therefore the primitive-level expression of the intrinsicality clause
	\textup{(SP1)}: a comparison is a relation among states, not a
	stipulated labeling of them.
\end{definition}

No topology, differentiable structure, metric, linear structure,
probability measure, or background arena is assumed at this level.
The admissibility stance of the monograph is strict:
all higher structure must be justified as internally determined by the
comparison system and its closed-world coherence conditions.

\section{Standing principle}
\begin{standingprinciple}[Closed-World Admissibility]
\label{sp:closed-world}
Throughout the monograph we work under one standing principle, unpacked
into the following clauses.
\begin{enumerate}[label=\textup{(SP\arabic*)}]
\item\label{sp:intrinsic}
      \textbf{Intrinsicality clause.}
      No background manifold, external observer frame, or auxiliary
      state-space primitive is admissible unless reconstructed internally
      from comparison data.
\item\label{sp:compactness}
      \textbf{Compactness clause.}
	  Finite-stage and global admissibility are linked by the exact
	  finite-to-global boundary of the refinement tower. Under
	  intrinsicality this boundary is finitary: by
	  \cref{lem:orbit-separation-finite-stage} every comparison separation is
	  witnessed at a single coordinate, so the refinement-tower limit requires
	  no completion and the finite-to-global boundary of
	  \cref{thm:boundary,thm:stability} holds without separate hypothesis. See
	  \cref{lem:orbit-separation-finite-stage,rem:no-limit-only-separation}.
\item\label{sp:closure}
	  \textbf{Comparison-completeness clause.}
	  Closedness is profile-pair completeness (rectangular completeness).
	  Under the intrinsicality condition in the comparison-world definition
	  and local
	  distinguishability, this is not an independent hypothesis but a
	  consequence; see \cref{thm:rectangular-completeness-intrinsicality}.
\item\label{sp:quotient}
	      \textbf{Quotient-descent clause.}
      Admissible semantic content is exactly the content descending through
      intrinsic quotient maps.
\item\label{sp:transport}
      \textbf{Transport-visibility clause.}
      Any admissible residue beyond quotient semantics is carried by the
      transport-obstruction hierarchy and its first visible layers.
\end{enumerate}
\end{standingprinciple}
\begin{remark}[How hypotheses are used]
In subsequent chapters, theorem headers use either
\emph{local setup hypotheses} (e.g., choosing objects, indices, or
connected components) or \emph{realization-level technical hypotheses}
(e.g., smooth realizability, jet injectivity). These are not independent
foundational axioms; they are scoped instantiations of
\cref{sp:intrinsic,sp:compactness,sp:closure,sp:quotient,sp:transport}.
Any non-local quantitative boundary condition left after theorem-level
discharge is recorded explicitly in \cref{app:conditional-completion}.
\end{remark}

\section{Methodological strategy}
The derivation is organized as a finite-to-global and algebra-to-geometry
program.
At finite stage, one controls compatibility through Boolean admissibility;
at global stage, one passes through exact compactness boundaries.
On the semantic side, comparison completeness determines canonical
quotient structure.
On the enrichment side, one proves that non-quotient content can occur
through at least one of two loci, possibly both, and no third.
On the transport side, one isolates a canonical filtration and proves
that its first stable visible obstruction is quadratic.
Under compatible smooth realization, the stabilized quadratic carrier
enters the intrinsic--smooth interface: the degree--\(2\)
placement is forced by triangle minimality, commutator filtration, and
Stone-limit stabilization, while faithful realization on that forced
carrier makes nonzero stabilized square classes equivalent to nonzero
realized curvature.  The faithfulness condition itself is then
decomposed and discharged: it is equivalent to the ghost-free regime
consisting of a torsion condition \textup{(T)}, a theorem in the Magnus
regime, and a single intrinsic detectability axiom \textup{(D)}, and
the tower supplies a canonical realization on which the second-jet
comparison map is injective by construction
(\cref{sec:interface-ghosts,sec:interface-detectability,sec:interface-canonical}).
The subsequent macroscopic, scalar, and
realization-level consequences proceed through that realized-channel
package.

At a conceptual level, the logical progression is
\[
		\begin{aligned}
			(U,\mathcal C)
			\;&\Longrightarrow\;\text{admissibility boundary and canonical quotient semantics}
			\\
			\;&\Longrightarrow\;\text{two-locus enrichment and transport filtration}
			\\
			\;&\Longrightarrow\;\text{first stable quadratic obstruction}
			\\
			\;&\Longrightarrow\;\text{intrinsic--smooth interface and macroscopic compatibility laws}
			\\
			\;&\Longrightarrow\;\text{scalar-channel classification, numerical closure, and canonical realization}
			\\
			\;&\Longrightarrow\;\text{state-side observable exhaustivity and the automorphism bridge}.
		\end{aligned}
	\]

\section{Main structural claim}
Theorem~\ref{thm:intro-structural-rigidity} records the structural
content proved across
\cref{ch:foundation,ch:bottom,ch:rotation,ch:closure,ch:locus,ch:lift,ch:triangles,ch:descent_obstruction,ch:bridge,ch:flow,ch:stitching,ch:christoffel,ch:interface,ch:causality,ch:hilbert,ch:einstein,ch:connection,ch:em,ch:matter,ch:matter-numbers,ch:numerical-closure,ch:matter-realization,ch:dynamics-program,app:fusion}.

\begin{theorem}[Structural rigidity of closed relational systems]
	\label{thm:intro-structural-rigidity}
	Let \((U,\mathcal C)\) be a comparison world satisfying the closed-system
	admissibility and comparison-completeness conditions developed in this
	monograph.
	Then the following structure is determined.
	\begin{enumerate}[label=\textup{(\arabic*)}]
		\item
		      Admissible state content factors through canonical quotient
		      semantics, with intrinsic product decomposition
		      \(U\cong X_A\times X_B\) and physical quotient
		      \(\Phys=(X_A\times X_B)/G\).
		\item
		      Beyond quotient semantics, admissible enrichment occurs through at
		      least one of two loci, possibly both---object-level representative
		      choice and morphism-level transport---with no third locus.
		\item
		      Transport admits a canonical descending filtration
		      \(F^1\supset F^2\supset F^3\supset\cdots\),
		      whose first stable visible obstruction is the quadratic layer
		      \(F^2/F^3\).
		\item
		      The same obstruction appears as finite triangle failure,
		      global loop-descent obstruction, and their bridge identification on
		      minimal subsystems.
		\item
		      The obstruction governs the internal observer-reduction framework and
		      the conditional regime of compatible irreversible descent, and admits
		      a canonical spacetime interleaving realization via a two-parameter
		      inverse-limit object.
		\item
		      Under smooth realization, the stabilized quadratic carrier admits an
		      intrinsic--smooth interface: the refinement tower is proved to
		      stabilize its first nontrivial transport obstruction at degree
		      \(2\), and faithful smooth realization on that forced carrier makes a
		      nonzero stabilized square class equivalent to nonzero realized
		      curvature on the realized degree--\(2\) channel.  The faithfulness
		      condition is itself decomposed into the torsion condition
		      \textup{(T)} and the detectability axiom \textup{(D)}, shown
		      necessary by explicit ghost counter-models, and discharged by a
		      canonical realization constructed from the tower.  This
		      degree--\(2\) realized-channel package feeds causal, Hilbert,
		      Einstein-boundary, connection, electromagnetic, matter-sector, and
		      numerical-closure consequences.
		\item
		      The scalar channel determined by the quadratic carrier is then
		      classified intrinsically, assigned its canonical matter-sector
		      realization role, and finally yields a concrete nonperturbative
		      invariant in the magnetic-confinement appendix.
		\item
		      In the separately marked state-side reconstruction extension,
		      relative to the additional realization data
		      \textup{(R1)}--\textup{(R3)}, the raw limit comparison
		      algebra is proved to be larger than every realized smooth
		      reading. The excess is exactly the fiber-separating sector.
		      Under the scale-coherence condition \textup{(U)}, the maximal
		      readable sector is identified with \(C(M)\), its smooth
		      derivations are identified with vector fields on \(M\), and
		      its transport-stabilizing derivations are identified with the
		      symmetry algebra of the realized connection up to gauge.
	\end{enumerate}
\end{theorem}
\begin{proof}
	Each item is proved in the indicated chapter block of the monograph.
	Item \textup{(1)} follows from the compactness-to-quotient chain in
	\cref{ch:foundation,ch:bottom,ch:rotation,ch:closure}.
	Item \textup{(2)} is the two-locus classification proved in
	\cref{ch:locus,ch:lift}.
	Items \textup{(3)} and \textup{(4)} are established by the transport
	obstruction chain in
	\cref{ch:triangles,ch:descent_obstruction,ch:bridge}.
		Item \textup{(5)} is derived in
		\cref{ch:flow,ch:stitching}.
		Item \textup{(6)} is proved through the smooth-realization chapter, the
		intrinsic-to-smooth interface, and their macroscopic
		continuations in
		\cref{ch:christoffel,ch:interface,ch:causality,ch:hilbert,ch:einstein,ch:connection,ch:em,ch:matter,ch:matter-numbers,ch:numerical-closure}.
		Item \textup{(7)} is completed by the scalar-classification and
		realization chain in
		\cref{ch:matter,ch:matter-numbers,ch:numerical-closure,ch:matter-realization,app:fusion}.
		Item \textup{(8)} is the state-side reconstruction extension proved in
		\cref{ch:dynamics-program}. It is stated relative to its own standing
		data and does not enter the dependency chain for items
		\textup{(1)}--\textup{(7)}.
		Thus the theorem is exactly the consolidated statement of the chapterwise
		results listed above.
\end{proof}

The point of the theorem is methodological as much as structural:
it does not append geometry, field theory, or matter-sector realization
to an independently chosen background.
It shows how these sectors arise as closure-compatible realizations of
structure already determined at the relational level, with scalar and
realization-level consequences forced downstream of the same
obstruction chain.

\section{Logical status of the construction}
The monograph proves theorem-level necessity results under the
explicit admissibility hypotheses developed in the text.
The dependency order is the reverse of the standard ordering
"geometry first, fields second":
semantic closure and transport obstruction first, then geometric,
field-theoretic, scalar, and realization-level structure, followed by a
separately marked state-side reconstruction extension and appendix-level
applications.

The foundational admissibility content of the monograph is concentrated
in a single primitive condition. By
\cref{lem:orbit-separation-finite-stage,lem:finite-support} and
\cref{thm:rectangular-completeness-intrinsicality}, the compactness
clause \textup{(SP2)} and the comparison-completeness clause
\textup{(SP3)} are consequences of intrinsicality
in the comparison-world definition together with local distinguishability.
Intrinsicality is an admissibility condition on primitive comparisons:
a predicate must be
invariant under the symmetries it helps constitute. The program therefore
rests on one identity---to be a distinction is to be a
symmetry-invariant of comparison---which is the intrinsicality clause
\textup{(SP1)} stated as a definition of the primitive rather than a
prohibition on later constructions. Later smooth, Hilbert, phase,
observer-side, and quantitative clauses enter as scoped realization data,
derived interface results, or explicitly recorded boundary conditions.

The resulting contribution is a sharpened criterion for admissibility in
closed-system modeling:
which constructions remain valid when all external reference structures
are removed.

\section{Organization of the proof}
\label{sec:paper}
The detailed reading paths are given in the Reader Guide.  The proof
itself has a single dependency order.
\Cref{ch:foundation,ch:bottom,ch:rotation,ch:closure} fix the
finite-to-global boundary, derive comparison completeness, and establish
quotient semantics.  \Cref{ch:locus,ch:lift,ch:triangles,ch:descent_obstruction,ch:bridge}
classify enrichment and isolate transport obstruction.  The middle
chapters then pass from obstruction to observer reduction, spacetime
interleaving, curvature, and the intrinsic--smooth interface.  The later
chapters treat the realized Hilbert, Einstein-boundary, connection,
electromagnetic, scalar, matter-sector, numerical, and state-side
consequences under their stated hypotheses.

Thus the chapter order records the order in which structures become
admissible.

\section{Conclusion}
\label{sec:introduction-conclusion}
This introduction fixes the mathematical entry point of the monograph:
primitive comparison data, closed-world admissibility, and the structural
theorem whose proof is distributed across the monograph.  The chapter
architecture is therefore a proof architecture, ordered from compactness
and quotient semantics through transport obstruction to geometric,
field-theoretic, scalar, realization-level, and state-side reconstruction
results.

Accordingly, \cref{ch:foundation} is logically first: it determines the
finite-to-global admissibility boundary that is reused by every later
construction.

\begin{chapterclosing}
\closingitem{Established}{The introduction fixes primitive comparison data, closed-world admissibility, and the dependency order of the proof.}
\closingitem{Carried forward}{The next step is the finite-to-global admissibility boundary on which comparison completeness depends.}
\end{chapterclosing}
