% !TeX root = closed-systems-from-comparison-completeness.tex
% ============================================================
\chapter{\texorpdfstring{Intrinsic \(2\)-Skeleton Obstruction on the Stone Limit}{Intrinsic 2-Skeleton Obstruction on the Stone Limit}}
\label{ch:interface}
% ============================================================
\section{Introduction}
Under the standing principle of closed-world admissibility
(\cref{sp:closed-world}), this chapter develops the inverse-limit
interface consequences of \cref{sp:compactness,sp:transport}.
Taking the triangle boundary and curvature realization from
\cref{ch:triangles,ch:christoffel} as fixed input, it proves that the
first stable intrinsic transport carrier is the quadratic carrier,
stabilizes that carrier on the Stone limit, and then relates nonzero
stabilized square classes to nonzero realized curvature for
\cref{ch:causality,ch:hilbert,ch:einstein} under faithful smooth
realization on this already-forced carrier.
\Cref{ch:triangles,ch:christoffel} established two structural facts
(\cref{thm:triangle,cor:first-visible-comm,thm:christoffel-main}).
First, the first finite witness of morphism-level transport defect is
triangular. Triangle relations therefore define the intrinsic
\(2\)-skeletal boundary of the transport system.
Second, the first visible graded carrier of intrinsic transport
noncommutativity is the quadratic quotient
\[
	F^2/F^3,
\]
and, under smooth realization, its smooth infinitesimal realization is
Riemann curvature.
The purpose of the present chapter is to connect these two facts
intrinsically, before any smooth realization is imposed.
The mechanism is algebraic.
At each finite stage, the isotropy group carries the augmentation
filtration
\[
	F^1K \supseteq F^2K \supseteq F^3K \supseteq \cdots,
\]
and commutator squares automatically land in the quadratic layer
\(F^2K\). Their images in
\[
	F^2K/F^3K
\]
define canonical degree--\(2\) transport classes.
The main point is that nontrivial triangle obstruction determines the
quadratic carrier, that this determination is forced rather than
postulated, that compatible quadratic square classes are preserved
under refinement and hence stabilize on the Stone limit, and that under
smooth realization faithful on this carrier a nonzero stabilized square
class is equivalent to nonzero realized curvature.
The chapter is organized as follows.
We first make the refinement-compatible transport system explicit at the
level of based loop groups and isotropy groups.
We then define the projective triangle obstruction.
Next we establish quadratic cancellation and define the degree--\(2\)
defect classes.
We then isolate the passage from triangle obstruction to the quadratic
carrier.
After that we prove that triangle-orientation reversal descends to an
intrinsic involution on the stabilized quadratic carrier.
Finally we establish, under smooth realization faithful on the forced
quadratic carrier, that a nonzero stabilized square class is equivalent
to nonzero realized curvature.
Thus the index ``\(2\)'' in the interface is not a hypothesis: it is
the theorem proved by the triangle minimality, commutator-filtration,
and refinement-stabilization chain.  The remaining faithfulness
condition belongs only to the realization map; it says that smooth
realization does not collapse the already forced degree--\(2\) carrier.
The closing sections of the chapter
(\cref{sec:interface-ghosts,sec:interface-detectability,sec:interface-canonical})
determine the exact status of that residual condition.  It decomposes
into a torsion condition \textup{(T)} on the carrier, which is a
theorem in the Magnus regime, and a detectability axiom \textup{(D)},
which is stated entirely on the tower; the pair is necessary, by
explicit ghost counter-models, and sufficient, by a canonical
realization constructed from the tower itself on which the second-jet
comparison map is injective by construction.  The faithfulness
condition is thereby discharged rather than assumed.
% ============================================================
\section{Refinement-compatible transport}
\label{sec:interface-refinement}
Let
\[
	\B_1\subseteq \B_2\subseteq \cdots
\]
be the Boolean refinement tower, and write
\[
	S_k:=\UF(\B_k),
	\qquad
	S_\infty \cong \varprojlim_k S_k
\]
for the associated Stone spaces.
For each stage \(k\), let
\[
	\mathcal N_k=(V_k,E_k)
\]
be a finite directed transport network equipped with a transport functor
\[
	\Hol_k:\Path^\pm(\mathcal N_k)\longrightarrow \mathsf C_k.
\]
Fix a basepoint \(p_k\in V_k\), and define
\[
	\Omega_{p_k}(\mathcal N_k)
	:=
	\Hom_{\Path^\pm(\mathcal N_k)}(p_k,p_k),
	\qquad
	K_{p_k}^{(k)}
	:=
	\Aut_{\mathsf C_k}(p_k).
\]
The restriction of \(\Hol_k\) to based loops is the defect map
\[
	\delta_{p_k}^{(k)}:
	\Omega_{p_k}(\mathcal N_k)\longrightarrow K_{p_k}^{(k)}.
\]
Assume the refinement package provides bonding functors
\[
	R_{k+1,k}:\Path^\pm(\mathcal N_{k+1})\to \Path^\pm(\mathcal N_k),
	\qquad
	Q_{k+1,k}:\mathsf C_{k+1}\to \mathsf C_k,
\]
such that
\begin{equation}\label{eq:interface-compat}
	Q_{k+1,k}\circ \Hol_{k+1}
	=
	\Hol_k\circ R_{k+1,k}.
\end{equation}
\begin{lemma}[Bonding maps on based loops and isotropy]
	\label{lem:interface-bonding}
	The refinement functors induce homomorphisms
	\[
		R_{k+1,k}^{\Omega}:
		\Omega_{p_{k+1}}(\mathcal N_{k+1})\longrightarrow
		\Omega_{p_k}(\mathcal N_k),
		\qquad
		Q_{k+1,k}^{K}:
		K_{p_{k+1}}^{(k+1)}\longrightarrow
		K_{p_k}^{(k)},
	\]
	and these satisfy
	\[
		Q_{k+1,k}^{K}\circ \delta_{p_{k+1}}^{(k+1)}
		=
		\delta_{p_k}^{(k)}\circ R_{k+1,k}^{\Omega}.
	\]
\end{lemma}
\begin{proof}
	Since \(R_{k+1,k}\) is a functor, it sends loops at \(p_{k+1}\) to loops
	at \(p_k\), hence restricts to a homomorphism
	\[
		R_{k+1,k}^{\Omega}:
		\Omega_{p_{k+1}}(\mathcal N_{k+1})\to \Omega_{p_k}(\mathcal N_k).
	\]
	Likewise \(Q_{k+1,k}\) restricts on isotropy groups to
	\[
		Q_{k+1,k}^{K}:
		\Aut_{\mathsf C_{k+1}}(p_{k+1})\to \Aut_{\mathsf C_k}(p_k).
	\]
	Now restrict \cref{eq:interface-compat} to
	\(\Omega_{p_{k+1}}(\mathcal N_{k+1})\). This yields
	\[
		Q_{k+1,k}^{K}\circ \delta_{p_{k+1}}^{(k+1)}
		=
		\delta_{p_k}^{(k)}\circ R_{k+1,k}^{\Omega},
	\]
	as claimed.
\end{proof}
% ============================================================
\section{Triangle obstruction}
\label{sec:interface-triangle}
For each stage \(k\), let
\[
	N_{\triangle,p_k}^{(k)}\trianglelefteq \Omega_{p_k}(\mathcal N_k)
\]
denote the normal closure of all based triangle loops.
\begin{definition}[Finite-stage triangle obstruction]
	Define
	\[
		\Obs_\triangle^{(k)}
		:=
		\delta_{p_k}^{(k)}\bigl(N_{\triangle,p_k}^{(k)}\bigr)
		\subseteq
		K_{p_k}^{(k)}.
	\]
\end{definition}
\begin{proposition}[Projective triangle obstruction]
	The groups \(\Obs_\triangle^{(k)}\) form a projective system under
	refinement. Consequently
	\[
		\Obs_\triangle^\infty
		:=
		\varprojlim_k \Obs_\triangle^{(k)}
	\]
	is canonically defined on the inverse-limit locus.
\end{proposition}
\begin{proof}
	Refinement preserves triangle incidence, hence sends a based triangle
	loop at stage \(k+1\) to a based triangle loop at stage \(k\). Since
	\(N_{\triangle,p_k}^{(k)}\) is the normal closure of the based triangle loops at stage
	\(k\), it follows that
	\[
		R_{k+1,k}^{\Omega}\bigl(N_{\triangle,p_{k+1}}^{(k+1)}\bigr)
		\subseteq
		N_{\triangle,p_k}^{(k)}.
	\]
	Applying \cref{lem:interface-bonding} gives
	\[
		Q_{k+1,k}^{K}\Bigl(\delta_{p_{k+1}}^{(k+1)}\bigl(N_{\triangle,p_{k+1}}^{(k+1)}\bigr)\Bigr)
		=
		\delta_{p_k}^{(k)}\Bigl(R_{k+1,k}^{\Omega}\bigl(N_{\triangle,p_{k+1}}^{(k+1)}\bigr)\Bigr)
		\subseteq
		\delta_{p_k}^{(k)}\bigl(N_{\triangle,p_k}^{(k)}\bigr).
	\]
	That is,
	\[
		Q_{k+1,k}^{K}\bigl(\Obs_\triangle^{(k+1)}\bigr)
		\subseteq
		\Obs_\triangle^{(k)}.
	\]
	Hence the triangle obstruction groups form a projective system.
\end{proof}
\begin{remark}[Minimal \(2\)-skeletal character]
	Triangle loops are the minimal \(2\)-cell boundaries in the intrinsic
	transport complex.
	Accordingly \(\Obs_\triangle^{(k)}\) is the first obstruction that is
	genuinely \(2\)-skeletal rather than merely path-theoretic.
\end{remark}
% ============================================================
\section{Augmentation filtration and quadratic cancellation}
\label{sec:interface-filtration}
Let \(K\) be a group and
\[
	\varepsilon:\mathbb Z[K]\to \mathbb Z
\]
its augmentation homomorphism. Write
\[
	I:=\ker(\varepsilon).
\]
\begin{definition}[Intrinsic augmentation filtration]
	For \(m\ge 1\), define
	\[
		F^mK
		:=
		\{g\in K: g-1\in I^m\}.
	\]
\end{definition}
\begin{lemma}[Commutators raise augmentation order]
	\label{lem:interface-comm-raises}
	For all integers \(r,s\ge 1\),
	\[
		[F^rK,F^sK]\subseteq F^{r+s}K.
	\]
\end{lemma}
\begin{proof}
	Let \(a\in F^rK\) and \(b\in F^sK\). Then
	\[
		a-1\in I^r,
		\qquad
		b-1\in I^s.
	\]
	Write
	\[
		u:=a-1\in I^r,
		\qquad
		v:=b-1\in I^s.
	\]
	Then
	\[
		a=1+u,
		\qquad
		b=1+v.
	\]
	In the \(I\)-adic completion one has convergent geometric-series
	expansions
	\[
		a^{-1}=1-u+u^2-u^3+\cdots,
		\qquad
		b^{-1}=1-v+v^2-v^3+\cdots.
	\]
	Hence
	\[
		a^{-1}-1\in I^r,
		\qquad
		b^{-1}-1\in I^s.
	\]
	Now compute
	\[
		[a,b]-1=aba^{-1}b^{-1}-1.
	\]
	Modulo \(I^{r+s}\), every nonconstant term in the product
	\[
		(1+u)(1+v)\bigl(1+(a^{-1}-1)\bigr)\bigl(1+(b^{-1}-1)\bigr)
	\]
	contains at least one factor from \(I^r\) and at least one factor from
	\(I^s\), or else lies in a strictly higher power of \(I\). Hence
	\[
		[a,b]-1\in I^{r+s}.
	\]
	Therefore
	\[
		[a,b]\in F^{r+s}K.
	\]
\end{proof}
Now fix a stage \(k\) and a basepoint \(p_k\).
For based loops \(e_i,e_j\in \Omega_{p_k}(\mathcal N_k)\), define the
commutator square
\[
	\square_{ij}:=e_ie_je_i^{-1}e_j^{-1}.
\]
\begin{proposition}[Quadratic cancellation]
	For every stage \(k\) and every commutator square \(\square_{ij}\),
	\[
		\delta_{p_k}^{(k)}(\square_{ij})\in F^2K_{p_k}^{(k)}.
	\]
\end{proposition}
\begin{proof}
	Since \(\delta_{p_k}^{(k)}\) is a homomorphism,
	\[
		\delta_{p_k}^{(k)}(\square_{ij})
		=
		[\delta_{p_k}^{(k)}(e_i),\delta_{p_k}^{(k)}(e_j)].
	\]
	Every group element lies in \(F^1K\). Hence
	\cref{lem:interface-comm-raises} with \(r=s=1\) gives
	\[
		\delta_{p_k}^{(k)}(\square_{ij})\in F^2K_{p_k}^{(k)}.
	\]
\end{proof}
\begin{lemma}[Primitive two-cell Magnus term]
\label{lem:primitive-two-cell-magnus}
Let \(a,b\in K\) be two transport germs whose degree--\(1\) classes
\[
\bar a,\bar b\in F^1K/F^2K
\]
are independent.  Then the commutator square
\[
[a,b]=aba^{-1}b^{-1}
\]
has nonzero quadratic class
\[
[a,b]\bmod F^3K\in F^2K/F^3K.
\]
More precisely, in the augmentation ideal one has
\[
[a,b]-1\equiv (a-1)(b-1)-(b-1)(a-1) \pmod {I^3},
\]
so the class of \([a,b]\) is the exterior product
\(\bar a\wedge\bar b\).
\end{lemma}
\begin{proof}
Write \(u=a-1\) and \(v=b-1\).  Modulo \(I^3\),
\[
	a^{-1}\equiv 1-u+u^2,
	\qquad
	b^{-1}\equiv 1-v+v^2.
\]
Multiplying
\[
	(1+u)(1+v)(1-u+u^2)(1-v+v^2)
\]
and discarding terms in \(I^3\) gives
\[
	[a,b]\equiv 1+uv-vu \pmod {I^3}.
\]
Thus
\[
	[a,b]-1\equiv uv-vu \pmod {I^3}.
\]
The antisymmetric tensor \(uv-vu\) is the image of
\(\bar a\wedge\bar b\).  Since \(\bar a\) and \(\bar b\) are independent,
this exterior product is nonzero.  Hence \([a,b]\notin F^3K\), while
\cref{lem:interface-comm-raises} gives \([a,b]\in F^2K\).
\end{proof}
\begin{lemma}[Minimal triangle witnesses have nonzero area class]
\label{lem:minimal-triangle-area}
A primitive triangle witness in the strict triangle regime determines two
first-order edge germs with nonzero exterior product in \(F^1K/F^2K\).
\end{lemma}
\begin{proof}
By \cref{thm:triangle} and \cref{cor:bridge-minimal}, every two-edge
restriction of a primitive triangle witness is coherent, while the full
three-edge subsystem is not.  If the degree--\(1\) edge germs of the
triangle spanned a rank--\(0\) or rank--\(1\) subgroup, then the full
triangle would factor through the same one-dimensional first-order
transport line as its two-edge restrictions.  In that case pairwise
coherence would determine the third edge and would extend to a coherent
vertex gauge on the full triangle, contradicting the strict minimality of
\cref{thm:triangle}.  Hence the primitive triangle witness has a rank--\(2\)
first-order span.  Choosing two independent edge germs \(a,b\) in that span
gives \(\bar a\wedge\bar b\neq0\).
\end{proof}
% ============================================================
\section{Quadratic defect classes}
\label{sec:interface-quadratic}
\begin{definition}[Quadratic defect class]
	For each stage \(k\), define
	\[
		\mathcal A_{ij}^{(k)}(p_k)
		:=
		\bigl[\delta_{p_k}^{(k)}(\square_{ij})\bigr]
		\in
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}.
	\]
\end{definition}
\begin{proposition}[Alternation and graded bilinearity]
	The assignment
	\[
		(i,j)\longmapsto \mathcal A_{ij}^{(k)}(p_k)
	\]
	is alternating, and commutator induces a bilinear map
	\[
		\bigl(F^1K_{p_k}^{(k)}/F^2K_{p_k}^{(k)}\bigr)
		\times
		\bigl(F^1K_{p_k}^{(k)}/F^2K_{p_k}^{(k)}\bigr)
		\longrightarrow
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}.
	\]
\end{proposition}
\begin{proof}
	By \cref{lem:interface-comm-raises}, the commutator of two degree--\(1\)
	elements lands in degree \(2\). Therefore the commutator descends to a
	map
	\[
		\bigl(F^1K/F^2K\bigr)\times\bigl(F^1K/F^2K\bigr)\to F^2K/F^3K.
	\]
	To prove bilinearity, let \(a,a',b\in F^1K\). The standard commutator
	identity gives
	\[
		[aa',b]=[a,b]\,[a,b,a']\,[a',b],
	\]
	where \([a,b,a']=[[a,b],a']\). Since \([a,b]\in F^2K\), one has
	\[
		[a,b,a']\in [F^2K,F^1K]\subseteq F^3K
	\]
	by \cref{lem:interface-comm-raises}. Hence modulo \(F^3K\),
	\[
		[aa',b]\equiv [a,b][a',b].
	\]
	Similarly,
	\[
		[a,bb']\equiv [a,b][a,b'] \pmod{F^3K}.
	\]
	Thus the induced map is bilinear on the quotients.
	For alternation, one has
	\[
		[a,a]=e
	\]
	for every \(a\), so the induced class of \([a,a]\) is zero in
	\(F^2K/F^3K\). Also,
	\[
		[b,a]=[a,b]^{-1},
	\]
	so in additive notation on the abelian quotient \(F^2K/F^3K\),
	\[
		\mathcal A_{ji}^{(k)}(p_k)=-\mathcal A_{ij}^{(k)}(p_k).
	\]
	Therefore the assignment is alternating.
\end{proof}
% ============================================================
\section{\texorpdfstring{Persistence of triangle obstruction in degree \(2\)}{Persistence of triangle obstruction in degree 2}}
\label{sec:interface-persistence}
The next step is the intrinsic passage from triangle obstruction to a
nonzero quadratic carrier.
\begin{lemma}[Primitive triangle defects survive quadratically]
	\label{lem:quadratic-nondegenerate}
	Let
	\[
		g\neq e
	\]
	be a primitive transport defect arising from a triangle witness.
	Then there exists a refinement stage \(\ell\) such that
	\[
		g\notin F^3K_{p_\ell}^{(\ell)}.
	\]
\end{lemma}
\begin{proof}
	By \cref{lem:minimal-triangle-area}, the primitive triangle witness
	determines two first-order edge germs \(a,b\) with
	\[
		\bar a\wedge\bar b\neq0.
	\]
	The comparison of the two opposite orderings of these germs is the
	associated commutator square.  By
	\cref{lem:primitive-two-cell-magnus}, that square has nonzero image in
	\[
		F^2K/F^3K.
	\]
	Refinement compatibility carries this primitive square class to the
	corresponding square class at sufficiently fine stages.  Therefore the
	refined primitive defect cannot lie in \(F^3\) at every stage.  Hence
	there exists \(\ell\) with
	\[
		g\notin F^3K_{p_\ell}^{(\ell)}.
	\]
\end{proof}
\begin{lemma}[Persistence of nontrivial triangle defect in \(F^2/F^3\)]
	\label{lem:interface-persistence}
	If
	\[
		\Obs_\triangle^{(k)}\neq \{e\},
	\]
	then, after passing if necessary to a sufficiently fine refinement stage
	\(\ell\ge k\), the image of \(\Obs_\triangle^{(k)}\) in
	\[
		F^2K_{p_\ell}^{(\ell)}/F^3K_{p_\ell}^{(\ell)}
	\]
	is nonzero.
\end{lemma}
\begin{proof}
	The group \(\Obs_\triangle^{(k)}\) is generated by the images of based
	primitive triangle loops and their conjugates.  If every primitive
	triangle defect had trivial image, the generated obstruction group would
	be trivial.  Since \(\Obs_\triangle^{(k)}\neq \{e\}\), at least one
	primitive triangle witness has nontrivial defect.

	By refinement compatibility, that primitive defect propagates to finer
	stages.  By \cref{lem:quadratic-nondegenerate}, there exists a refinement
	stage \(\ell\ge k\) at which its refined image does not lie in
	\[
		F^3K_{p_\ell}^{(\ell)}.
	\]
	On the other hand, the witness is a boundary-supported \(2\)-cell
	defect, so its degree--\(1\) image vanishes by
	\cref{cor:triangle-degree1-invisible} in \cref{ch:christoffel}.  Its
	first possible nonzero image is therefore in the quadratic range.  Hence
	its class in
	\[
		F^2K_{p_\ell}^{(\ell)}/F^3K_{p_\ell}^{(\ell)}
	\]
	is nonzero, and so the image of \(\Obs_\triangle^{(k)}\) in this quotient
	is nonzero.
\end{proof}
% ============================================================
\section{The relevant quadratic carrier}
\label{sec:interface-generation}
The next object is the intrinsic degree--\(2\) carrier generated by
quadratic square classes coming from the triangle-induced transport
sector.
\begin{definition}[Finite-stage quadratic carrier]
	\label{def:interface-Kk}
	Let
	\[
		\mathcal K_k
		\subseteq
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}
	\]
	denote the subgroup generated by the classes
	\[
		\mathcal A_{ij}^{(k)}(p_k)
	\]
	arising from commutator squares in the triangle-closed transport
	regime.
\end{definition}
\begin{lemma}[Quadratic carrier is square-generated]
	For each stage \(k\), the subgroup
	\[
		\mathcal K_k
		\subseteq
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}
	\]
	defined in \cref{def:interface-Kk} is generated by the classes
	\[
		\mathcal A_{ij}^{(k)}(p_k).
	\]
\end{lemma}
\begin{proof}
	This is immediate from the definition of \(\mathcal K_k\).
\end{proof}
\begin{remark}[What is retained from triangle defect]
	The carrier \(\mathcal K_k\) is the intrinsic degree--\(2\) residue of
	the \(2\)-skeletal transport boundary.
	It is exactly the portion of triangle obstruction that remains visible
	after quotienting away cubic and higher transport error.
\end{remark}
% ============================================================
\section{Quadratic carrier determined by triangle obstruction}
\label{sec:interface-carrier}
\begin{theorem}[Triangle obstruction determines a quadratic carrier]
	\label{thm:interface-quad-carrier}
	If
	\[
		\Obs_\triangle^{(k)}\neq \{e\},
	\]
	then after refinement there exist \(\ell\ge k\) and indices
	\((i,j)\) such that
	\[
		\mathcal A_{ij}^{(\ell)}(p_\ell)\neq 0.
	\]
\end{theorem}
\begin{proof}
	By \cref{lem:interface-persistence}, after refinement to some
	\(\ell\ge k\), the image of \(\Obs_\triangle^{(k)}\) in
	\[
		F^2K_{p_\ell}^{(\ell)}/F^3K_{p_\ell}^{(\ell)}
	\]
	is nonzero.
	By definition, the relevant degree--\(2\) carrier is the subgroup
	\(\mathcal K_\ell\) generated by the square classes
	\[
		\mathcal A_{ij}^{(\ell)}(p_\ell).
	\]
	Therefore at least one such class must be nonzero.
\end{proof}
% ============================================================
\section{Universal property of the quadratic carrier}
\label{sec:interface-universal}
\Cref{thm:interface-quad-carrier} shows that the degree--\(2\) quotient is the first
place where intrinsic \(2\)-skeletal obstruction becomes visible.
The next proposition isolates its universal property.
\begin{proposition}[Universal property of the quadratic carrier]
	\label{prop:interface-universal-quadratic}
	Let
	\[
		\pi_{2,3}:F^2K_p\to F^2K_p/F^3K_p
	\]
	be the quotient map.
	Then \(\pi_{2,3}\) is initial among group morphisms from \(F^2K_p\) to abelian
	groups which annihilate \(F^3K_p\).
	Equivalently, if
	\[
		\phi:F^2K_p\to A
	\]
	is a group morphism into an abelian group \(A\) with
	\[
		F^3K_p\subseteq \ker(\phi),
	\]
	then there exists a unique morphism
	\[
		\widetilde\phi:F^2K_p/F^3K_p\to A
	\]
	such that
	\[
		\phi=\widetilde\phi\circ \pi_{2,3}.
	\]
\end{proposition}
\begin{proof}
	This is the universal property of the quotient group
	\(F^2K_p/F^3K_p\).
\end{proof}
\begin{remark}
	\Cref{prop:interface-universal-quadratic} gives the precise sense in
	which \(F^2/F^3\) is the canonical carrier of quadratic transport
	defect.
	Any detector defined on the quadratic layer and insensitive to cubic
	error factors uniquely factors through this quotient.
\end{remark}
% ============================================================
\section{Stabilization on the Stone limit}
\label{sec:interface-stabilization}
\begin{definition}[Stabilized quadratic carrier]
	Define
	\[
		\mathcal K_\infty
		:=
		\varprojlim_k \mathcal K_k.
	\]
\end{definition}
\begin{theorem}[Stabilized quadratic defect]
	\label{thm:interface-stabilized}
	Every compatible family of finite-stage quadratic square classes defines
	an element of
	\[
		\mathcal K_\infty=\varprojlim_k \mathcal K_k.
	\]
	Conversely, every element of \(\mathcal K_\infty\) is represented by a
	compatible family whose \(k\)-th component lies in \(\mathcal K_k\).
\end{theorem}
\begin{proof}
	By \cref{lem:interface-bonding}, the refinement package supplies
	group homomorphisms
	\(Q_{k+1,k}^{K}:K_{p_{k+1}}^{(k+1)}\to K_{p_k}^{(k)}\). Any group
	homomorphism extends linearly to a morphism of group rings preserving
	augmentation ideals, so \(Q_{k+1,k}^{K}\) carries
	\(F^2K_{p_{k+1}}^{(k+1)}\) into \(F^2K_{p_k}^{(k)}\) and
	\(F^3K_{p_{k+1}}^{(k+1)}\) into \(F^3K_{p_k}^{(k)}\). Therefore it
	induces a homomorphism
	\(F^2K_{p_{k+1}}^{(k+1)}/F^3K_{p_{k+1}}^{(k+1)}\to
	F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}\). Because each
	\(\mathcal K_k\) is generated by the square classes
	\(\mathcal A_{ij}^{(k)}(p_k)\) and refinement carries such square
	classes to square classes at the previous stage, these quotient maps
	restrict to bonding maps \(\mathcal K_{k+1}\to\mathcal K_k\). By
	definition of inverse limit, compatible families of finite-stage
	square classes then determine exactly the elements of
	\(\mathcal K_\infty=\varprojlim_k \mathcal K_k\), and every element
	of \(\mathcal K_\infty\) is represented by such a compatible family.
\end{proof}
\begin{definition}[Stabilized square class]
	Whenever a compatible family of finite-stage square classes
	\[
		\bigl(\mathcal A_{ij}^{(k)}(p_k)\bigr)_k
	\]
	is fixed, we write its image in \(\mathcal K_\infty\) as
	\[
		\mathcal A_{ij}(p_\infty).
	\]
\end{definition}
\begin{theorem}[Second-layer stabilization theorem]
	\label{thm:interface-seal}
	In the closed comparison stack, the first stable nontrivial transport
	obstruction in the refinement tower is the stabilized quadratic carrier
	\[
		\mathcal K_\infty\simeq F^2/F^3.
	\]
	Equivalently, a nontrivial triangle obstruction cannot first appear in
	degree \(1\), and it cannot be supported purely in \(F^3\) or in any
	higher filtration layer.  After refinement it has a nonzero image in the
	degree--\(2\) carrier, and compatible such images stabilize on the Stone
	limit.
\end{theorem}
\begin{proof}
	By \cref{thm:triangle} and \cref{cor:bridge-minimal}, the first finite
	morphism-level obstruction is triangular: every two-edge restriction is
	coherent, while the full three-edge subsystem need not be.  Hence no
	degree--\(1\) transport quotient can carry this obstruction.  This is the
	content of \cref{cor:triangle-degree1-invisible} in
	\cref{ch:christoffel}: boundary-supported \(2\)-cell defect has trivial
	image in \(F^1/F^2\).

	By \cref{lem:interface-persistence}, any nontrivial triangle obstruction
	has, after sufficiently fine refinement, a nonzero image in
	\[
		F^2K_{p_\ell}^{(\ell)}/F^3K_{p_\ell}^{(\ell)}.
	\]
	By \cref{thm:interface-quad-carrier}, this nonzero image is represented
	by a nonzero quadratic square class
	\[
		\mathcal A_{ij}^{(\ell)}(p_\ell).
	\]
	Equivalently, the obstruction cannot first appear in degree \(3\) or
	higher: such an appearance would force the degree--\(2\) image to vanish,
	contradicting the nonzero square class just obtained.

	Finally, \cref{thm:interface-stabilized} shows that refinement-compatible
	families of these finite-stage square classes define exactly the inverse
	limit carrier
	\[
		\mathcal K_\infty=\varprojlim_k\mathcal K_k.
	\]
	Thus the first stable nontrivial transport obstruction of the refinement
	tower is neither degree \(1\) nor a hidden higher layer.  It is exactly
	the stabilized quadratic carrier \(\mathcal K_\infty\simeq F^2/F^3\).
\end{proof}
\begin{remark}[What has stabilized]
	The object \(\mathcal K_\infty\) is the intrinsic degree--\(2\)
	transport carrier retained by the full refinement tower.
	It is the first stabilized obstruction visible on the Stone inverse
	limit.
\end{remark}
\begin{definition}[Admissible degree--\(2\) scalar detector]
	\label{def:interface-admissible-scalar}
	Fix a basepoint \(p\). A group morphism
	\[
		\phi:F^2K_p\to \mathbb R
	\]
	is called an admissible degree--\(2\) scalar detector if it is invariant
	under multiplication by elements of \(F^3K_p\), i.e. if
	\[
		\phi(xz)=\phi(x)
		\qquad
		\text{for all }x\in F^2K_p,\ z\in F^3K_p.
	\]
\end{definition}
\begin{theorem}[Kernel annihilation at the quadratic carrier]
	\label{thm:kernel-annihilation-upstream}
	Fix a basepoint \(p\), and let
	\[
		\pi_{2,3}:F^2K_p\to F^2K_p/F^3K_p
	\]
	be the canonical quotient.
	Let
	\[
		\phi:F^2K_p\to \mathbb R
	\]
	be any admissible degree--\(2\) scalar detector.
	Then
	\[
		F^3K_p\subseteq \Ker(\phi),
	\]
	and therefore there exists a unique morphism
	\[
		\overline\phi:F^2K_p/F^3K_p\to\mathbb R
	\]
	such that
	\[
		\phi=\overline\phi\circ \pi_{2,3}.
	\]
\end{theorem}
\begin{proof}
	Let \(z\in F^3K_p\).
	Since \(\phi\) is admissible, \cref{def:interface-admissible-scalar}
	gives
	\[
		\phi(xz)=\phi(x)
		\qquad
		\text{for all }x\in F^2K_p.
	\]
	Taking \(x=e\), where \(e\) is the identity element of \(F^2K_p\),
	yields
	\[
		\phi(z)=\phi(e).
	\]
	Because \(\phi\) is a group morphism into the additive group
	\(\mathbb R\), one has
	\[
		\phi(e)=0.
	\]
	Hence
	\[
		\phi(z)=0
		\qquad
		\text{for all }z\in F^3K_p.
	\]
	Therefore
	\[
		F^3K_p\subseteq \Ker(\phi).
	\]
	Now apply \cref{prop:interface-universal-quadratic}. Since \(\phi\)
	annihilates \(F^3K_p\), there exists a unique morphism
	\[
		\overline\phi:F^2K_p/F^3K_p\to\mathbb R
	\]
	such that
	\[
		\phi=\overline\phi\circ \pi_{2,3}.
	\]
	This is the required factorization.
\end{proof}
% ============================================================
\section{Orientation reversal on the quadratic carrier}
\label{sec:orientation-carrier}
Primitive triangle witnesses admit a natural orientation reversal.
We show that this operation descends to the stabilized quadratic carrier.
\begin{definition}[Primitive orientation reversal]
	Let
	\[
		\omega(p,q,r)
	\]
	denote the defect class associated to the oriented triangle
	\((p,q,r)\).
	Define the reversed class by
	\[
		\mathsf R_\triangle\,\omega(p,q,r):=\omega(p,r,q).
	\]
\end{definition}
\begin{lemma}[Orientation reversal preserves the triangle quotient]
	Let \(w_\triangle(p,q,r)\) denote the triangle boundary word.
	Then reversing orientation sends
	\[
		w_\triangle(p,q,r)\mapsto w_\triangle(p,r,q)=w_\triangle(p,q,r)^{-1}.
	\]
	Hence orientation reversal preserves the normal closure of triangle
	relations.
\end{lemma}
\begin{proof}
	By definition
	\[
		w_\triangle(p,q,r)=[rp][qr][pq].
	\]
	Reversing orientation exchanges the last two vertices and yields the
	inverse word.
	Since the triangle relations are imposed through their normal closure,
	the quotient group is invariant under inversion.
\end{proof}
\begin{theorem}[Orientation involution on the quadratic carrier]
	Primitive triangle reversal descends to a well-defined involution
	\[
		\mathsf R_{\mathrm{or}}:\mathcal K_\infty\longrightarrow \mathcal K_\infty.
	\]
	Moreover
	\[
		\mathsf R_{\mathrm{or}}^2=I.
	\]
\end{theorem}
\begin{proof}
	At each finite stage, the quotient
	\[
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)}
	\]
	is abelian, so inversion is a well-defined automorphism.
	On square classes it exchanges
	\[
		\mathcal A_{ij}^{(k)}(p_k)
		\longleftrightarrow
		\mathcal A_{ji}^{(k)}(p_k)
		=
		-\mathcal A_{ij}^{(k)}(p_k).
	\]
	By refinement compatibility, these involutions are compatible from stage
	to stage and therefore descend to the inverse limit
	\(\mathcal K_\infty\).
	Applying inversion twice yields the identity, so
	\[
		\mathsf R_{\mathrm{or}}^2=I.
	\]
\end{proof}
\begin{remark}
	The involution \(\mathsf R_{\mathrm{or}}\) exchanges the two orientations of every primitive
	triangle witness.
	Later chapters may use the corresponding \(\pm1\)-eigenspace
	decomposition to organize the defect sector.
\end{remark}
% ============================================================
\section{Smooth realization}
\label{sec:interface-smooth}
Assume now a smooth realization in which square-loop transport admits a
small-loop parameter \(\varepsilon\), and suppose parallel transport
around a coordinate square satisfies
\[
	\Hol(\square_{ij}(\varepsilon))
	=
	\id
	+
	\varepsilon^2 R(\partial_i,\partial_j)
	+
	O(\varepsilon^3).
\]
Fix a limit basepoint \(p_\infty\in S_\infty\), and set
\[
	V:=T_{\iota(p_\infty)}M.
\]
\begin{definition}[Second-jet comparison map]
	\label{def:interface-J2}
	Define
	\[
		\Jtwo:\mathcal K_\infty\to \End(V)
	\]
	to be the map sending the quadratic class of an intrinsic square defect
	to the coefficient of \(\varepsilon^2\) in the realized holonomy
	expansion.
\end{definition}
\begin{lemma}[Well-definedness of the second-jet map]
	The map \(\Jtwo\) is well-defined on \(\mathcal K_\infty\).
\end{lemma}
\begin{proof}
	By \cref{lem:filtration-holonomy-order} in \cref{ch:christoffel}, the
	quotient
	\[
		F^2/F^3
	\]
	records exactly the quadratic holonomy coefficient, while terms of order
	\(3\) and higher are killed by passage to the quotient by \(F^3\).
	Hence two representatives of the same stabilized quadratic class have
	the same \(\varepsilon^2\)-coefficient under smooth realization.
	Therefore \(\Jtwo\) is well-defined.
\end{proof}
\begin{theorem}[Curvature identification]
	\label{thm:interface-curvature}
	For the stabilized quadratic class,
	\[
		\Jtwo\bigl(\mathcal A_{ij}(p_\infty)\bigr)
		=
		R(\partial_i,\partial_j).
	\]
\end{theorem}
\begin{proof}
	By definition,
	\[
		\mathcal A_{ij}(p_\infty)
	\]
	is the degree--\(2\) class of the intrinsic commutator defect of the
	square loop.
	By \cref{def:interface-J2}, \(\Jtwo\) extracts the coefficient of
	\(\varepsilon^2\) in the realized holonomy expansion.
	By \cref{cor:infdef} in \cref{ch:christoffel}, the coefficient of
	\(\varepsilon^2\) in the holonomy expansion of a small square is
	exactly
	\[
		R(\partial_i,\partial_j).
	\]
	Therefore
	\[
		\Jtwo\bigl(\mathcal A_{ij}(p_\infty)\bigr)
		=
		R(\partial_i,\partial_j).
	\]
\end{proof}
% ============================================================
\section{Interface theorem}
\label{sec:interface-main}
\begin{theorem}[Intrinsic--smooth interface]
	\label{thm:interface-main}
	Under the standing principle \cref{sp:closed-world}, the refinement
	tower stabilizes its first nontrivial transport obstruction at the
	degree--\(2\) carrier
	\[
		\mathcal K_\infty\simeq F^2/F^3.
	\]
	Assume only realization faithfulness on this forced carrier, i.e. that
	the comparison map
	\[
		\Jtwo:\mathcal K_\infty\to \End(V)
	\]
	is injective.
	Then the following are equivalent:
	\[
		\exists\,\mathcal A_{ij}(p_\infty)\neq0,
		\qquad
		R\neq0.
	\]
\end{theorem}
\begin{proof}
	The stabilization assertion is \cref{thm:interface-seal}.  Thus the
	appearance of the second layer is not an additional interface assumption;
	it is the intrinsic output of the refinement tower.  What remains to
	check is the faithful smooth reading of that carrier.

	Suppose that
	\[
		\mathcal A_{ij}(p_\infty)\neq 0.
	\]
	Since \(\Jtwo\) is injective,
	\[
		\Jtwo\bigl(\mathcal A_{ij}(p_\infty)\bigr)\neq 0.
	\]
	By \cref{thm:interface-curvature},
	\[
		\Jtwo\bigl(\mathcal A_{ij}(p_\infty)\bigr)
		=
		R(\partial_i,\partial_j),
	\]
	hence
	\[
		R\neq 0.
	\]
	Conversely, if
	\[
		R\neq 0,
	\]
	then for some indices \(i,j\),
	\[
		R(\partial_i,\partial_j)\neq 0.
	\]
	By \cref{thm:interface-curvature},
	\[
		\Jtwo\bigl(\mathcal A_{ij}(p_\infty)\bigr)=R(\partial_i,\partial_j)\neq 0.
	\]
	Since the zero class has zero second-jet coefficient by
	\cref{def:interface-J2}, this implies
	\[
		\mathcal A_{ij}(p_\infty)\neq 0.
	\]
\end{proof}
\begin{remark}[Status of second-layer stabilization]
\label{rem:second-jet-status}
The aggressive question is not whether one may assume the second layer;
the answer is that one may not.  The second layer is forced.  The proof is
the chain
\[
\begin{aligned}
&\text{strict triangle minimality}
\Longrightarrow
\text{no degree--1 residue}
\\
&\Longrightarrow
\text{commutator-square defect in }F^2
\Longrightarrow
\text{nonzero class in }F^2/F^3
\\
&\Longrightarrow
\mathcal K_\infty.
\end{aligned}
\]
The only remaining hypothesis is not that the tower stabilizes at degree
\(2\), but that a chosen compatible smooth realization is faithful on the
already stabilized degree--\(2\) carrier.  Under that faithful reading,
\cref{thm:interface-curvature} identifies the realized image with Riemann
curvature, and the GR chain proceeds.

\Cref{sec:interface-ghosts,sec:interface-detectability,sec:interface-canonical}
settle the status of this last hypothesis.  Faithfulness decomposes
exactly into the torsion condition \textup{(T)} of
\cref{def:condition-T} and the detectability axiom \textup{(D)} of
\cref{def:axiom-D}: each half is defeated by an explicit ghost
counter-model when dropped, so the decomposition is not refinable; and
under \textup{(T)} and \textup{(D)} together, the canonical realization
of \cref{thm:canonical-realization} satisfies the faithfulness
condition by construction.  What was an interface assumption about an
external smooth world is thereby converted into one intrinsic
non-degeneracy axiom on the tower, together with a torsion computation
that is a theorem in the Magnus regime.
\end{remark}
% ============================================================
\section{Ghost sectors and the limits of forcing}
\label{sec:interface-ghosts}
The interface theorem leaves exactly one hypothesis in play: that a
chosen compatible smooth realization is faithful on the already forced
degree--\(2\) carrier.  The present section determines how much of that
hypothesis can be a theorem.  The answer is structural.  Faithfulness
can fail in precisely two ways, called here \emph{ghost sectors}: a
torsion ghost, in which a nonzero stabilized class has finite order and
is therefore invisible to every real-linear reading; and a disconnected
ghost, in which the realized transport completion is totally
disconnected and admits no differential reading at all.  Both failure
modes are realized by explicit counter-models, so no theorem of the
closed stack alone can force faithfulness.  What can be forced is the
exact decomposition of the condition into the two corresponding
non-degeneracy statements, one algebraic and one topological, and the
construction of a canonical realization on which faithfulness holds by
construction once those statements are in force.

Throughout these closing sections, write
\[
	G_\infty:=\varprojlim_k K_{p_k}^{(k)}
\]
for the limit isotropy group along the bonding maps
\(Q_{k+1,k}^{K}\) of \cref{lem:interface-bonding}.
\begin{definition}[Second-jet faithfulness condition]
	\label{def:second-jet-faithfulness}
	A compatible smooth realization is \emph{faithful on the stabilized
		quadratic carrier} if its second-jet comparison map
	\[
		\Jtwo:\mathcal K_\infty\to \End(V)
	\]
	of \cref{def:interface-J2} is injective.  This is the condition
	inherited by the subsequent chapters.
\end{definition}
\begin{lemma}[Finite generation at each stage]
	\label{lem:transport-fg}
	At each stage \(k\), the based loop group
	\(\Omega_{p_k}(\mathcal N_k)\) is a finitely generated free group, and
	consequently the subgroup
	\(\mathsf H_k\le F^1K_{p_k}^{(k)}/F^2K_{p_k}^{(k)}\)
	generated by the degree--\(1\) classes of the stage edge germs, and
	the stage carrier \(\mathcal K_k\), are finitely generated abelian
	groups.
\end{lemma}
\begin{proof}
	The network \(\mathcal N_k\) is finite, and
	\(\Path^\pm(\mathcal N_k)\) is the free groupoid on its edges.  The
	vertex group of a free groupoid on a finite graph is free of finite
	rank, by the spanning-tree presentation of the based loop group.
	The defect map \(\delta_{p_k}^{(k)}\) is a homomorphism, so the
	degree--\(1\) classes of edge germs generate a finitely generated
	subgroup \(\mathsf H_k\) of the abelian group \(F^1K/F^2K\), and the
	finitely many square classes \(\mathcal A_{ij}^{(k)}(p_k)\) generate
	the finitely generated abelian group \(\mathcal K_k\).
\end{proof}
\begin{lemma}[Real readings factor through the rationalized carrier]
	\label{lem:real-factorization}
	Let
	\[
		\phi:\mathcal K_\infty\to W
	\]
	be any homomorphism of abelian groups into the additive group of a
	vector space \(W\) over \(\mathbb Q\) or \(\mathbb R\).  Then \(\phi\)
	factors uniquely as
	\[
		\phi=\phi_{\mathbb Q}\circ\iota_{\mathbb Q},
		\qquad
		\iota_{\mathbb Q}:\mathcal K_\infty\to\mathcal K_\infty\otimes_{\mathbb Z}\mathbb Q,
	\]
	with \(\phi_{\mathbb Q}\) \(\mathbb Q\)-linear.  Moreover
	\[
		\ker(\iota_{\mathbb Q})
		=
		\operatorname{Tor}(\mathcal K_\infty),
	\]
	the torsion subgroup of the carrier.
\end{lemma}
\begin{proof}
	The additive group of \(W\) is uniquely divisible, so
	\[
		\phi_{\mathbb Q}\Bigl(\kappa\otimes\tfrac{m}{n}\Bigr)
		:=
		\tfrac{m}{n}\,\phi(\kappa)
	\]
	is well defined and \(\mathbb Q\)-linear, and it is the unique
	\(\mathbb Q\)-linear map satisfying
	\(\phi_{\mathbb Q}\circ\iota_{\mathbb Q}=\phi\), since simple tensors
	span the rationalization.
	For the kernel, write \(\mathbb Q\) as the direct limit of the chain
	\(\mathbb Z\xrightarrow{\,2\,}\mathbb Z\xrightarrow{\,3\,}\mathbb
	Z\xrightarrow{\,4\,}\cdots\).  Tensoring with \(\mathcal K_\infty\)
	exhibits \(\mathcal K_\infty\otimes\mathbb Q\) as the direct limit of
	the chain of multiplication maps on \(\mathcal K_\infty\), so
	\(\iota_{\mathbb Q}(\kappa)=0\) if and only if some finite composite
	of multiplications kills \(\kappa\), i.e.\ if and only if
	\(n\kappa=0\) for some \(n\ge1\).
\end{proof}
\begin{corollary}[Torsion ghosts are invisible]
	\label{cor:torsion-invisible}
	For every compatible smooth realization,
	\[
		\operatorname{Tor}(\mathcal K_\infty)\subseteq\ker(\Jtwo).
	\]
	In particular, if the carrier has a nonzero torsion class, then no
	compatible smooth realization whatsoever is faithful.  Faithfulness
	therefore requires that \(\mathcal K_\infty\) be torsion-free, and
	this requirement is not an artifact of any realization choice.
\end{corollary}
\begin{proof}
	The comparison map \(\Jtwo\) is a homomorphism of abelian groups into
	the additive group of the real vector space \(\End(V)\): holonomy of
	a product of square defects multiplies, and modulo \(F^3\) the
	quadratic coefficients add, as in the proof of well-definedness of
	\(\Jtwo\).  Apply \cref{lem:real-factorization}.
\end{proof}
\begin{definition}[Condition \textup{(T)}: torsion-freeness of the carrier]
	\label{def:condition-T}
	The closed tower satisfies \emph{condition \textup{(T)}} if the
	stabilized quadratic carrier \(\mathcal K_\infty\) is torsion-free.
\end{definition}
\begin{definition}[Magnus regime]
	\label{def:magnus-regime}
	A stage \(k\) is in the \emph{Magnus regime} if \(\mathsf H_k\) is
	torsion-free and the graded pairing of
	\cref{lem:primitive-two-cell-magnus}, extended by the alternation and
	bilinearity of the commutator pairing, embeds the exterior square:
	that is, the natural map
	\[
		\mu_k:\Lambda^2\mathsf H_k\longrightarrow
		F^2K_{p_k}^{(k)}/F^3K_{p_k}^{(k)},
		\qquad
		\bar a\wedge\bar b\longmapsto[a,b]\bmod F^3K,
	\]
	is injective.
\end{definition}
\begin{theorem}[Condition \textup{(T)} is a theorem in the Magnus regime]
	\label{thm:magnus-torsion-free}
	Suppose cofinally many stabilized stages are in the Magnus regime.
	Then every stage carrier \(\mathcal K_k\) at such a stage is a
	finitely generated free abelian group, and
	\[
		\mathcal K_\infty=\varprojlim_k\mathcal K_k
	\]
	is torsion-free.  Condition \textup{(T)} holds.
\end{theorem}
\begin{proof}
	Fix a stage \(k\) in the Magnus regime.  By construction
	\(\mathcal K_k\) is generated by the square classes
	\(\mathcal A_{ij}^{(k)}(p_k)\), and by
	\cref{lem:primitive-two-cell-magnus} each square class is the image
	\(\mu_k(\bar e_i\wedge\bar e_j)\) of a wedge of degree--\(1\) edge
	classes.  Hence
	\[
		\mathcal K_k\subseteq\operatorname{im}(\mu_k)\cong\Lambda^2\mathsf H_k .
	\]
	By \cref{lem:transport-fg}, \(\mathsf H_k\) is finitely generated,
	and in the Magnus regime it is torsion-free, hence free abelian of
	finite rank; so \(\Lambda^2\mathsf H_k\) is free abelian of finite
	rank, and the subgroup \(\mathcal K_k\) is finitely generated free
	abelian, in particular torsion-free.
	Finally, the inverse limit embeds in the product
	\(\prod_k\mathcal K_k\); a torsion element of the limit would be a
	torsion element in each coordinate, and each coordinate is
	torsion-free at cofinally many stages.  Hence
	\(\mathcal K_\infty\) is torsion-free.
\end{proof}
\begin{remark}[Scope of the Magnus regime]
	\label{rem:magnus-scope}
	The Magnus regime is not an exotic hypothesis; it is the regime in
	which the degree--\(2\) graded algebra of transport words behaves as
	in the free model.  For free transport groups it is a theorem of
	Magnus and Witt: the augmentation-graded ring of a free group ring is
	the tensor ring on the abelianization, in which the antisymmetric
	tensors \(uv-vu\) span a direct summand isomorphic to the exterior
	square, so \(\mu_k\) is injective \cite{Magnus1935,Witt1937}.
	Labute's determination of the
	graded Lie ring of one-relator presentations, and the
	Duchamp--Krob analysis of partially commutative
	(right-angled Artin) presentations, extend the same graded control to
	the natural presentation classes in which triangle-relation stage
	groups live \cite{Labute1970,DuchampKrob1992}.  \Cref{lem:minimal-triangle-area} already supplies the
	germ-level input: primitive triangle witnesses have rank--\(2\)
	first-order span, hence nonzero wedge.  Outside the Magnus regime,
	condition \textup{(T)} is exactly the statement that the tower has no
	torsion ghosts, and \cref{cor:torsion-invisible} shows it is the
	necessary half of faithfulness, not a convenience.
\end{remark}
\begin{remark}[Disconnected ghosts: the tower cannot force its own
	realization]
	\label{rem:disconnected-ghost}
	The same intrinsic tower admits compatible realizations with no
	differential content.  Compose each stage transport functor with
	reduction to a tower of finite quotients of the isotropy groups; the
	limit transport then lands in a profinite group: compact, totally
	disconnected, with a neighborhood basis of open subgroups at the
	identity.  Every identity neighborhood of such a group contains
	nontrivial subgroups, and a totally disconnected group admits no
	nonconstant continuous family \(g(\varepsilon)\) of holonomies,
	since the continuous image of an interval is connected, hence a
	point.  In such a realization no small-loop expansion exists, and
	every differential reading of the carrier vanishes identically,
	while the entire intrinsic algebra of this chapter --- filtration,
	square classes, stabilization --- is untouched.  Consequently the
	selection of a smooth reading is genuinely additional structure:
	what the closed stack can force is the dichotomy of
	\cref{thm:structure-trichotomy} below, and the choice of branch is
	the single residual axiom, stated as \cref{def:axiom-D}.
\end{remark}
% ============================================================
\section{Detectability and the structure trichotomy}
\label{sec:interface-detectability}
The disconnected ghost of \cref{rem:disconnected-ghost} is excluded not
by smoothness assumptions but by a non-degeneracy statement about
defects themselves: a nonzero failure of closure should not be
infinitesimally small.  This section states that axiom and proves that,
once it is in force, the structure theory of locally compact groups
forces the Lie alternative.
\begin{definition}[Scale-compatible completion]
	\label{def:scale-completion}
	A \emph{scale-compatible completion} of the limit transport law is a
	homomorphism
	\[
		\tau:G_\infty\to G
	\]
	into a locally compact topological group \(G\), with dense image,
	together with a scale assignment \(\varepsilon_k\downarrow 0\) for
	the stages, such that for every refinement-compatible square family
	the realized defects
	\(\tau\bigl(\delta_{p_k}^{(k)}(\square_{ij})\bigr)\) admit a
	continuous interpolation
	\[
		g_{ij}:[0,\varepsilon_0)\to G,
		\qquad
		g_{ij}(\varepsilon_k)=\tau\bigl(\delta_{p_k}^{(k)}(\square_{ij})\bigr),
		\qquad
		g_{ij}(0)=\id .
	\]
	This is the minimal regularity recording that refinement is a scale
	limit.
\end{definition}
\begin{definition}[Axiom \textup{(D)}: uniform defect detectability]
	\label{def:axiom-D}
	A scale-compatible completion is \emph{detectable} if \(G\) has no
	small subgroups: some identity neighborhood \(U\subseteq G\) contains
	no nontrivial subgroup of \(G\).  The closed tower satisfies
	\emph{axiom \textup{(D)}} if its realized transport completion is
	detectable.
\end{definition}
\begin{remark}[Intrinsic reading of axiom \textup{(D)}]
	\label{rem:axiom-D-intrinsic}
	Detectability is a closure-native statement: there is a definite
	threshold below which the only defect arbitrarily close to closure is
	closure itself.  Equivalently, no nonzero stabilized class has
	realized representatives entering every identity neighborhood at
	every scale; failures of closure are uniformly detectable.  At the
	scalar level this is the archimedean axiom for defect magnitudes, and
	by H\"older's theorem an archimedean ordered abelian group of
	magnitudes order-embeds in \((\mathbb R,+)\) \cite{Holder1901}.  This is why the
	admissible degree--\(2\) scalar detectors of
	\cref{def:interface-admissible-scalar} are real-valued by necessity
	rather than by convention.  Axiom \textup{(D)} is the unique point of
	these closing sections at which structure beyond the intrinsic tower
	is invoked, and \cref{rem:disconnected-ghost} shows it is independent
	of the tower axioms: the profinite realization satisfies every
	intrinsic axiom and violates \textup{(D)}.
\end{remark}
\begin{theorem}[Structure trichotomy for completed transport]
	\label{thm:structure-trichotomy}
	Let \(\tau:G_\infty\to G\) be a scale-compatible completion.  Then
	exactly one of the following holds.
	\begin{enumerate}[label=\textup{(\alph*)}]
		\item
		      \emph{Lie branch.}  \(G\) has no small subgroups.  Then \(G\) is a
		      Lie group, by the Gleason--Yamabe theorem
		      \cite{Gleason1952,Yamabe1953,MontgomeryZippin1955,Tao2014}.
		      If moreover some
		      stabilized class realizes nontrivially, \(G\) is non-discrete, the
		      exponential map is a local homeomorphism onto an identity
		      neighborhood, every realized square family admits the quadratic
		      small-loop expansion, and the degree--\(2\) reading of the carrier
		      is real-linear: by \cref{lem:real-factorization} it factors
		      through \(\mathcal K_\infty\otimes\mathbb R\).
		\item
		      \emph{Disconnected branch.}  \(G\) is totally disconnected and
		      non-discrete.  Then \(G\) has a neighborhood basis of compact open
		      subgroups at the identity, by van Dantzig's theorem
		      \cite{vanDantzig1936}; every
		      identity neighborhood contains nontrivial subgroups; every
		      continuous interpolation \(g_{ij}\) is constant, so all
		      sufficiently fine realized defects are trivial and the
		      degree--\(2\) reading vanishes identically.
		\item
		      \emph{Mixed branch.}  \(G\) has small subgroups and nontrivial
		      identity component.  Then, by the Gleason--Yamabe theorem, every
		      identity neighborhood \(U\) contains a compact subgroup
		      \(N\subseteq U\), normal in an open subgroup \(G'\le G\), with
		      \(G'/N\) a Lie group; the degree--\(2\) reading factors through
		      these Lie quotients and annihilates exactly the classes realized
		      inside the ghost kernels \(N\).
	\end{enumerate}
	In every branch, a Lie reading of the carrier exists after the ghost
	sector is annihilated; the reading is faithful precisely when the
	ghost sector is trivial; and axiom \textup{(D)} holds precisely on
	branch \textup{(a)}.
\end{theorem}
\begin{proof}
	The three cases are exclusive and exhaustive: either \(G\) has no
	small subgroups, which is case \textup{(a)}; or it has small
	subgroups, and then it is either totally disconnected, case
	\textup{(b)} (the discrete possibility having no small subgroups and
	belonging to \textup{(a)}), or has nontrivial identity component,
	case \textup{(c)}.

	\textup{(a)}  A locally compact group with no small subgroups is a
	Lie group; this is the Gleason--Yamabe solution of Hilbert's fifth
	problem.  If \(G\) were discrete, every continuous interpolation
	\(g_{ij}\) with \(g_{ij}(0)=\id\) would be constant near \(0\),
	forcing all sufficiently fine realized defects to be trivial,
	contrary to the assumption that some stabilized class realizes
	nontrivially; hence \(G\) is a Lie group of positive dimension, the
	exponential map is a local homeomorphism, and in exponential
	coordinates each interpolation \(g_{ij}(\varepsilon)\) with
	\(g_{ij}(0)=\id\) has a quadratic leading term on commutator-square
	families, because the degree--\(1\) terms cancel for closed loops
	exactly as in \cref{lem:primitive-two-cell-magnus}.  The coefficient
	assignment is additive on the carrier, hence by
	\cref{lem:real-factorization} factors through
	\(\mathcal K_\infty\otimes\mathbb R\).

	\textup{(b)}  A locally compact totally disconnected group has a
	neighborhood basis of compact open subgroups at the identity, by van
	Dantzig's theorem; if \(G\) is non-discrete these subgroups are
	nontrivial, so \(G\) has small subgroups.  A continuous map from the
	connected interval \([0,\varepsilon_0)\) into a totally disconnected
	space has connected image, hence is constant; with
	\(g_{ij}(0)=\id\) this forces
	\(\tau\bigl(\delta(\square_{ij})\bigr)=\id\) for all sufficiently
	fine stages, so every degree--\(2\) reading vanishes.

	\textup{(c)}  The Gleason--Yamabe theorem in its general form
	provides, inside every identity neighborhood \(U\), a compact
	subgroup \(N\subseteq U\) normal in an open subgroup \(G'\), with
	\(G'/N\) Lie.  Composing the reading with the quotient maps
	\(G'\to G'/N\) yields readings through Lie groups as in
	\textup{(a)}; a class whose realized representatives lie in the
	kernels \(N\) for every \(U\) is annihilated by all such readings,
	and conversely.

	The final statements collect \textup{(a)}--\textup{(c)}: in each
	branch the reading factors through a Lie group after annihilating
	the classes supported in ghost kernels (all classes, in branch
	\textup{(b)}); faithfulness is precisely triviality of what is
	annihilated; and no small subgroups is the defining condition of
	branch \textup{(a)}.
\end{proof}
\begin{remark}[Why the group and not an action]
	\label{rem:hilbert-smith}
	The trichotomy is run on the transport group itself, where the
	locally compact structure theory is complete.  The corresponding
	question for group \emph{actions} on a presupposed topological
	manifold is the Hilbert--Smith conjecture, which remains open in
	general and is settled only under additional regularity (Lipschitz
	actions, by Repov\v{s} and \v{S}\v{c}epin) or in low dimension
	(dimension three, by Pardon) \cite{RepovsScepin1997,Pardon2013}.  The connection-first order of the
	closed stack is therefore not a stylistic choice: it places the
	forcing question where it has a complete answer.
\end{remark}
% ============================================================
\section{The canonical realization: faithfulness by construction}
\label{sec:interface-canonical}
Under condition \textup{(T)}, the tower itself supplies a smooth
realization, with no external choices: the Malcev completion of its
two-step transport quotient.  On this canonical realization the
second-jet comparison map is injective by construction, so the
faithfulness condition of \cref{def:second-jet-faithfulness} is
discharged rather than assumed.
\begin{definition}[Canonical two-step model]
	\label{def:canonical-model}
	Assume condition \textup{(T)}, and fix a stabilized stage \(k\) in
	the regime of \cref{thm:interface-seal}.  Set
	\[
		\mathsf H_k^{\mathbb R}:=\mathsf H_k\otimes\mathbb R,
		\qquad
		\mathsf Z_k^{\mathbb R}:=\mathcal K_k\otimes\mathbb R,
	\]
	and let
	\[
		\mathfrak n_2^{(k)}
		:=
		\mathsf H_k^{\mathbb R}\oplus\mathsf Z_k^{\mathbb R}
	\]
	carry the bracket determined by the graded commutator pairing:
	\[
		[u+z,\;u'+z']
		:=
		\mu_k(u\wedge u'),
	\]
	extended \(\mathbb R\)-bilinearly, with \(\mathsf Z_k^{\mathbb R}\)
	central.  All double brackets vanish, so
	\(\mathfrak n_2^{(k)}\) is a two-step nilpotent Lie algebra.  The
	\emph{canonical two-step model} is the group
	\[
		\mathsf N_2^{(k)}
		:=
		\bigl(\mathfrak n_2^{(k)},\;\cdot\bigr),
		\qquad
		x\cdot y:=x+y+\tfrac12[x,y],
	\]
	the simply connected two-step nilpotent Lie group in exponential
	coordinates; this is the Malcev completion of the two-step transport
	quotient \cite{Malcev1949}.  The bonding maps \(Q_{k+1,k}^{K}\) induce graded Lie
	homomorphisms of the models, and on the stabilized carrier the
	degree--\(2\) layer of the limit model is
	\(\mathcal K_\infty\otimes\mathbb R\).
\end{definition}
\begin{lemma}[Exact square holonomy in two-step groups]
	\label{lem:two-step-holonomy}
	In \(\mathsf N_2^{(k)}\), for all \(u,v\in\mathsf H_k^{\mathbb R}\)
	and all \(\varepsilon\),
	\[
		(\varepsilon u)\cdot(\varepsilon v)\cdot(-\varepsilon u)\cdot(-\varepsilon v)
		=
		\varepsilon^2[u,v],
	\]
	with no error term.
\end{lemma}
\begin{proof}
	Compute with the terminating Baker--Campbell--Hausdorff law
	\(x\cdot y=x+y+\tfrac12[x,y]\), using centrality of brackets
	\cite{KobayashiNomizu1963}:
	\[
		(\varepsilon u)\cdot(\varepsilon v)
		=
		\varepsilon u+\varepsilon v+\tfrac12\varepsilon^2[u,v].
	\]
	Multiplying by \(-\varepsilon u\) gives
	\[
		\varepsilon v+\tfrac12\varepsilon^2[u,v]
		+\tfrac12[\varepsilon u+\varepsilon v,\,-\varepsilon u]
		=
		\varepsilon v+\varepsilon^2[u,v],
	\]
	and multiplying by \(-\varepsilon v\) gives
	\[
		\varepsilon^2[u,v]
		+\tfrac12[\varepsilon v,\,-\varepsilon v]
		=
		\varepsilon^2[u,v].
	\]
\end{proof}
\begin{lemma}[Faithful unipotent fiber model]
	\label{lem:unipotent-fiber}
	Let
	\[
		W_k:=\mathbb R\oplus\mathsf H_k^{\mathbb R}\oplus\mathsf Z_k^{\mathbb R},
	\]
	and for \(g=u+z\in\mathsf N_2^{(k)}\) define
	\[
		\rho(g)(t,x,w)
		:=
		\bigl(t,\;x+tu,\;w+tz+\tfrac12[u,x]\bigr).
	\]
	Then \(\rho\) is a faithful unipotent representation
	\(\rho:\mathsf N_2^{(k)}\to\GL(W_k)\).
\end{lemma}
\begin{proof}
	For \(g_i=u_i+z_i\),
	\[
		\rho(g_1)\rho(g_2)(t,x,w)
		=
		\bigl(t,\;x+t(u_1+u_2),\;
		w+t(z_1+z_2)+\tfrac12[u_1+u_2,x]+\tfrac12 t[u_1,u_2]\bigr),
	\]
	which equals \(\rho(g_1\cdot g_2)(t,x,w)\), since
	\(g_1\cdot g_2=(u_1+u_2)+(z_1+z_2+\tfrac12[u_1,u_2])\).
	If \(\rho(g)=\id\), evaluating at \((1,0,0)\) gives \(u=0\) and then
	\(z=0\); so \(\rho\) is faithful.  Finally \(\rho(g)-\id\) maps
	\(\mathbb R\) into \(\mathsf H\oplus\mathsf Z\), maps \(\mathsf H\)
	into \(\mathsf Z\), and kills \(\mathsf Z\); hence
	\((\rho(g)-\id)^3=0\) and \(\rho\) is unipotent.
\end{proof}
\begin{theorem}[Canonical realization; faithfulness by construction]
	\label{thm:canonical-realization}
	Assume condition \textup{(T)}.  Then the canonical two-step model,
	with fiber \(W_k\) and transport \(\rho\), is a compatible smooth
	realization of the stabilized transport law in the sense of
	\cref{sec:interface-smooth}: square-loop transport admits the
	small-loop parameter \(\varepsilon\), with
	\[
		\Hol(\square_{ij}(\varepsilon))
		=
		\id
		+
		\varepsilon^2\,\mathsf E\bigl([u_i,u_j]\bigr),
	\]
	exactly, where
	\(\mathsf E:\mathsf Z_k^{\mathbb R}\to\End(W_k)\),
	\(\mathsf E(z)(t,x,w):=(0,0,tz)\), is injective and real-linear.
	The second-jet comparison map of this realization is
	\[
		\Jtwo^{\mathrm{can}}
		=
		\mathsf E\circ\bigl(\kappa\mapsto\kappa\otimes1\bigr)
		:\mathcal K_\infty\to\End(W),
	\]
	whose kernel is \(\operatorname{Tor}(\mathcal K_\infty)=0\).  Hence
	\(\Jtwo^{\mathrm{can}}\) is injective: the second-jet faithfulness
	condition of \cref{def:second-jet-faithfulness} holds for the
	canonical realization by construction, and by
	\cref{thm:interface-main} nonzero stabilized square classes are
	equivalent to nonzero realized curvature there, with no interface
	assumption remaining.
\end{theorem}
\begin{proof}
	By \cref{lem:two-step-holonomy}, the realized defect of the
	commutator square at scale \(\varepsilon\) is the central element
	\(\varepsilon^2[u_i,u_j]\), and by \cref{lem:unipotent-fiber} its
	action on the fiber is
	\[
		\rho\bigl(\varepsilon^2[u_i,u_j]\bigr)(t,x,w)
		=
		\bigl(t,\;x,\;w+t\varepsilon^2[u_i,u_j]\bigr)
		=
		\bigl(\id+\varepsilon^2\,\mathsf E([u_i,u_j])\bigr)(t,x,w),
	\]
	with vanishing error term, which is the small-loop expansion of
	\cref{sec:interface-smooth} in exact form.  The
	\(\varepsilon^2\)-coefficient of the class
	\(\mathcal A_{ij}(p_\infty)\) is therefore
	\(\mathsf E\) applied to the image of the class in
	\(\mathcal K_\infty\otimes\mathbb R\) (stabilization,
	\cref{thm:interface-seal}, identifies the stabilized class with its
	stage representatives).  The map \(\kappa\mapsto\kappa\otimes1\) has
	kernel \(\operatorname{Tor}(\mathcal K_\infty)\) by
	\cref{lem:real-factorization}, which vanishes under condition
	\textup{(T)}; and \(\mathsf E\) is injective since
	\(\mathsf E(z)=0\) forces \(z=0\) at \(t=1\).  Hence
	\(\Jtwo^{\mathrm{can}}\) is injective.
\end{proof}
\begin{proposition}[Universality of the canonical reading]
	\label{prop:canonical-universality}
	Let any compatible smooth realization with comparison map
	\(\Jtwo:\mathcal K_\infty\to\End(V)\) be given.  Then there is a
	unique real-linear map
	\[
		L:\mathcal K_\infty\otimes\mathbb R\to\End(V)
	\]
	with
	\(\Jtwo=L\circ(\kappa\mapsto\kappa\otimes1)\).
	Consequently every realization's reading factors through the
	degree--\(2\) layer of the canonical model, and an arbitrary
	compatible smooth realization is faithful if and only if its
	comparison \(L\) is injective on the image of the canonical carrier.
\end{proposition}
\begin{proof}
	By \cref{lem:real-factorization}, \(\Jtwo\) factors uniquely
	\(\mathbb Q\)-linearly through
	\(\mathcal K_\infty\otimes\mathbb Q\), and a \(\mathbb Q\)-linear map
	into a real vector space extends uniquely \(\mathbb R\)-linearly to
	the real scalar extension
	\(\mathcal K_\infty\otimes\mathbb R
	=(\mathcal K_\infty\otimes\mathbb Q)\otimes_{\mathbb Q}\mathbb R\)
	by \(v\otimes r\mapsto r\,L_{\mathbb Q}(v)\).  The factorization
	statement and the faithfulness criterion follow, since
	\(\kappa\mapsto\kappa\otimes1\) is injective under condition
	\textup{(T)} and realizes the canonical carrier.
\end{proof}
\begin{proposition}[Equivariant rigidity of readings]
	\label{prop:equivariant-rigidity}
	Suppose the real carrier \(\mathcal K_\infty\otimes\mathbb R\)
	decomposes under the intrinsic stage symmetries, including the
	orientation involution \(\mathsf R_{\mathrm{or}}\), into channels on
	which the symmetry action is irreducible, and let a compatible smooth
	realization be equivariant for these symmetries.  Then on each
	channel the comparison \(L\) of \cref{prop:canonical-universality} is
	either zero or injective; and if the realization is leading-order
	exact on a channel, i.e.\ some class of the channel has nonzero
	realized quadratic coefficient, then \(L\) is injective on that
	channel.
\end{proposition}
\begin{proof}
	The kernel of \(L\) restricted to a channel is a symmetry-invariant
	subspace; by irreducibility it is zero or the whole channel.
	Leading-order exactness excludes the second alternative.
\end{proof}
\begin{remark}[External certifications]
	\label{rem:external-certifications}
	The placement and the discharge are corroborated by independent
	classical results.  The floor: in Riemann normal coordinates the
	\(0\)- and \(1\)-jet of a metric are gauge-trivial at a point, and
	the \(2\)-jet modulo diffeomorphisms is exactly the curvature
	tensor; for closed loops the degree--\(1\) holonomy term cancels
	identically, so the first invariant obstruction is quadratic ---
	the smooth shadow of \cref{thm:interface-seal}
	\cite{KobayashiNomizu1963,Wald1984}.  The cap: local
	operators have locally finite order (Peetre), natural bundles have
	finite order (Epstein--Thurston), and in four dimensions the only
	second-order metric field equations are the Einstein equations with
	cosmological constant (Lovelock); nothing above the second jet can
	enter the realized laws \cite{Peetre1959,EpsteinThurston1979,Lovelock1971}.
	The combinatorial-to-smooth reading has a
	concrete ancestor: the deficit angles of piecewise flat spaces
	converge to the smooth curvature measures
	(Cheeger--M\"uller--Schrader), which is the metric-space form of the
	assignment \(\mathcal A_{ij}\mapsto R(\partial_i,\partial_j)\)
	\cite{CheegerMullerSchrader1984}.  On
	the dynamical side, closure of the hypersurface-deformation algebra
	forces the Einstein Hamiltonian
	(Hojman--Kucha\v{r}--Teitelboim): path-independence of evolution,
	a closure axiom, already forces the dynamics once the kinematical
	arena is in place \cite{HojmanKucharTeitelboim1976}.  Standard general relativity, by contrast,
	postulates the smooth manifold outright, and the discrete-to-continuum
	programs (causal sets and their Hauptvermutung) meet exactly the
	interface treated here.
\end{remark}
\begin{remark}[Ghost-free reading of the subsequent chapters]
	\label{rem:ghost-free-reading}
	From this point on, the phrase ``under the inherited second-jet
	faithfulness condition'' may be read as ``in the ghost-free regime
	\textup{(T)}+\textup{(D)}.''  Condition \textup{(T)} is a theorem in
	the Magnus regime (\cref{thm:magnus-torsion-free}) and is necessary
	for faithfulness in any realization
	(\cref{cor:torsion-invisible}); axiom \textup{(D)} is the single
	intrinsic non-degeneracy axiom selecting the Lie branch of
	\cref{thm:structure-trichotomy}, and is independent of the tower
	axioms (\cref{rem:disconnected-ghost}); and under the pair, the
	canonical realization of \cref{thm:canonical-realization} satisfies
	the faithfulness condition by construction, while
	\cref{prop:canonical-universality,prop:equivariant-rigidity} transfer
	faithfulness to arbitrary equivariant, leading-order-exact
	realizations.  Nothing about an external smooth world is assumed
	anywhere in this chain.
\end{remark}
% ============================================================
\section{Einstein boundary}
\label{sec:interface-boundary}
The present chapter identifies the stabilized quadratic square-class
channel as the forced intrinsic degree--\(2\) transport carrier and,
under faithful smooth realization on that carrier, shows that nonzero
such classes are equivalent to nonzero realized curvature.
The faithfulness condition itself has been decomposed into the torsion
condition \textup{(T)} and the detectability axiom \textup{(D)}, shown
to be exactly the ghost-free regime, and discharged on the canonical
realization (\cref{thm:canonical-realization}).
No dynamical equations have been assumed or derived.
Macroscopic compatibility laws belong to the subsequent chapter.

\section{Conclusion}
\label{sec:interface-conclusion}
The interface result is now closed in stabilized form: the refinement
tower has a forced degree--\(2\) first obstruction, and under faithful
smooth realization on that carrier the existence of a nonzero stabilized
quadratic square class is equivalent to nonzero realized curvature on the
same limit object.  The residual faithfulness condition is no longer an
assumption about an external smooth world: it is the ghost-free regime
\textup{(T)}+\textup{(D)}, with \textup{(T)} a theorem in the Magnus
regime, \textup{(D)} a single intrinsic detectability axiom, and the
realization supplied canonically by the tower itself
(\cref{sec:interface-canonical}).

Accordingly, \cref{ch:causality} combines the stabilized quadratic
carrier, under the faithful-realization criterion that nonzero stabilized
square classes are equivalent to nonzero realized curvature, with the
observer-side filtration inherited from
\cref{ch:flow} and interleaved with the refinement tower in
\cref{ch:stitching} to derive the macroscopic causal and scaling
consequences forced by the closed stack.
