\documentclass[12pt,openany]{book}
% ============================================================
% Fonts and encoding
% ============================================================
\usepackage[T1]{fontenc}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{newunicodechar}
% ============================================================
% Page geometry
% ============================================================
\usepackage[margin=1.15in]{geometry}
% ============================================================
% Math packages
% ============================================================
\usepackage{amsmath,amssymb,mathtools,amsthm,mathrsfs}
\numberwithin{equation}{section}
% ============================================================
% Graphics
% ============================================================
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{tikz-cd}
% ============================================================
% Lists
% ============================================================
\usepackage{enumitem}
% ============================================================
% Hyperlinks and cross-references
% ============================================================
\usepackage[unicode,hypertexnames=false]{hyperref}
\ifdefined\pdfinfoomitdate
  \pdfinfoomitdate=1
\fi
\ifdefined\pdftrailerid
  \pdftrailerid{}
\fi
\ifdefined\pdfsuppressptexinfo
  \pdfsuppressptexinfo=15
\fi
\hypersetup{
  pdftitle  = {Closed Systems from Comparison Completeness: A Unified Theory},
  pdfauthor = {Chast K. Wolfe},
  pdfsubject = {Mathematical monograph on closed relational systems, comparison completeness, quotient semantics, transport, curvature, and electromagnetism},
  pdfkeywords = {closed systems, comparison completeness, relational systems, quotient semantics, holonomy, curvature, electromagnetism},
  pdfcreator = {LaTeX},
  pdfproducer = {LaTeX},
  pdflang = {en-US},
  pdfdisplaydoctitle = true,
  hidelinks
}
\usepackage[nameinlink,noabbrev]{cleveref}
% ============================================================
% Theorem environments
% ============================================================
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{standingprinciple}{Standing Principle}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\crefname{theorem}{Theorem}{Theorems}
\crefname{lemma}{Lemma}{Lemmas}
\crefname{definition}{Definition}{Definitions}
\crefname{remark}{Remark}{Remarks}
\crefname{corollary}{Corollary}{Corollaries}
\crefname{proposition}{Proposition}{Propositions}
\crefname{standingprinciple}{Standing Principle}{Standing Principles}
\renewcommand{\qedsymbol}{$\square$}
% ============================================================
% Unified macros
% ============================================================
\newcommand{\U}{\mathsf U}
\newcommand{\BA}{\mathsf B_A}
\newcommand{\BB}{\mathsf B_B}
\newcommand{\B}{\mathsf B}
\newcommand{\Phys}{\mathsf{Phys}}
\newcommand{\PhysH}{\mathsf{Phys}_{\mathcal{H}}}
\newcommand{\J}{\mathsf J}
\newcommand{\raw}{\mathrm{raw}}
\DeclareMathOperator{\UF}{UF}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Bij}{Bij}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Fix}{Fix}
\DeclareMathOperator{\Stab}{Stab}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\piZero}{\pi_0}
\DeclareMathOperator{\Idl}{Idl}
\DeclareMathOperator{\Ric}{Ric}
\DeclareMathOperator{\Scal}{Scal}
\DeclareMathOperator{\Diff}{Diff}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\dimdiamond}{dim_{\diamond}}
\DeclareMathOperator{\Riem}{Riem}
\DeclareMathOperator{\Ad}{Ad}
\DeclareMathOperator{\Obs}{Obs}
\DeclareMathOperator{\Homeo}{Homeo}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\Map}{Map}
\DeclareMathOperator{\Iso}{Iso}
\DeclareMathOperator{\Chan}{Chan}
\DeclareMathOperator{\Obj}{Ob}
\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\Bool}{Bool}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Vol}{Vol}
\DeclareMathOperator{\gr}{gr}
\DeclareMathOperator{\ab}{ab}
\DeclareMathOperator{\Curv}{Curv}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\diag}{diag}
\newcommand{\Hol}{\mathrm{Hol}}
\newcommand{\Path}{\mathrm{Path}}
\newcommand{\e}{e}
\newcommand{\acts}{\curvearrowright}
\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\disc}{\mathrm{disc}}
\newcommand{\GI}{\mathcal G(\mathcal I)}
\newcommand{\T}{\mathrm{T}}
\newcommand{\GL}{\mathrm{GL}}
\newcommand{\LC}{\mathrm{LC}}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\tri}{\triangle}
\newcommand{\Ob}{\mathrm{Ob}}
\newcommand{\Mor}{\mathrm{Mor}}
\newcommand{\Lie}{\mathcal L}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\End}{\mathrm{End}}
\newcommand{\owedge}{\mathbin{\wedge\!\!\!\!\wedge}}
\newcommand{\trtr}{\mathrm{tr}^{(2)}}
\newcommand{\ncl}[1]{\Bigl\langle\!\Bigl\langle\,#1\,\Bigr\rangle\!\Bigr\rangle}
\newcommand{\Jtwo}{\mathsf{Jet}_2}
\newcommand{\G}{\mathsf G}
\newcommand{\reg}{\mathrm{reg}}
\newunicodechar{∎}{\ensuremath{\square}}
\makeatletter
\let\ps@headings\ps@plain
\let\ps@myheadings\ps@plain
\makeatother
\hbadness=10000
\vbadness=10000
\hfuzz=\maxdimen
\vfuzz=\maxdimen
\pdfstringdefDisableCommands{%
  \def\Phys{Phys}%
  \def\PhysH{Phys\_H}%
  \def\Hol{Hol}%
  \def\Path{Path}%
  \def\B{B}%
}
% ============================================================
\begin{document}
\pagestyle{plain}
\sloppy
\frontmatter
\begin{titlepage}
	\centering
	\vspace*{3cm}
	{\Huge\bfseries Closed Systems from\\[6pt] Comparison Completeness}\\[1.5cm]
	{\Large\itshape A Theory of Closed Relational Systems}\\[10pt]
	{\Large Chast K.~Wolfe}\\[0.55cm]
	{\large Closure Research Initiative, United States}\\[4pt]
	{\normalsize \href{mailto:chast.wolfe@closureresearchinitiative.org}{\texttt{chast.wolfe@closureresearchinitiative.org}}}\\[0.6cm]
	{\normalsize Version 7, June 12, 2026}\\[2cm]
	\vfill
\end{titlepage}
\input{preface}
\tableofcontents
\mainmatter
\pagestyle{plain}
\input{ch00_introduction}
\part{Foundations}
\input{ch01_foundation}
\input{ch02_bottom}
\part{Quotient Semantics and the No-Go Theorem}
\input{ch03_rotation}
\input{ch04_closure}
\part{Lifting, Transport, and Holonomy}
\input{ch05_locus}
\input{ch06_lift}
\input{ch07_triangles}
\input{ch08_descent_obstruction}
\input{ch09_bridge}
\part{Dynamics, Irreversibility, and Spacetime}
\input{ch10_flow}
\input{ch11_stitching}
\part{Curvature and the Einstein Boundary}
\input{ch12_christoffel}
\input{ch13_interface}
\input{ch14_causality}
\input{ch15_hilbert}
\input{ch16_einstein}
\input{ch17_manifold}
\part{Electromagnetism}
\input{ch18_em}
\input{ch19_matter}
\part{Numerical Closure}
\input{ch20_numbers}
\input{ch21_shell_closure}
\input{ch22_matter_realization}
\part{Observable Exhaustivity and Dynamics}
\input{ch23_dynamics_program}
\appendix
\part{Appendix}
\input{app_conditional_completion}
\input{app_fusion}
\backmatter
\begin{thebibliography}{99}
	\bibitem{Isham1993}
	C.~J.~Isham,
	``Canonical Quantum Gravity and the Problem of Time,''
	in \textit{Integrable Systems, Quantum Groups, and Quantum Field Theories},
	ed.\ L.~A.~Ibort and M.~A.~Rodr\'{\i}guez
	(Kluwer, 1993), pp.~157--287;
	arXiv:gr-qc/9210011.
	\bibitem{Kuchar1992}
	K.~V.~Kucha\v{r},
	``Time and Interpretations of Quantum Gravity,''
	\textit{Int.\ J.\ Mod.\ Phys.\ D} \textbf{20} (2011), 3--86;
	based on the 1991 lectures published in
	\textit{Proc.\ 4th Canadian Conf.\ on General Relativity},
	ed.\ G.~Kunstatter et al.\ (World Scientific, 1992).
	DOI: \href{https://doi.org/10.1142/S0218271811019347}{10.1142/S0218271811019347}.
	\bibitem{Anderson2017}
	E.~Anderson,
	\emph{The Problem of Time: Quantum Mechanics Versus General Relativity}
	(Springer, 2017).
	DOI: \href{https://doi.org/10.1007/978-3-319-58848-3}{10.1007/978-3-319-58848-3}.
	\bibitem{Stone1936}
	M.~H.~Stone,
	``The theory of representations for Boolean algebras,''
	Trans.\ Amer.\ Math.\ Soc.\ \textbf{40} (1936), no.~1, 37--111.
	DOI: \href{https://doi.org/10.2307/1989664}{10.2307/1989664}.
	\bibitem{MacLane1998}
	S.~Mac~Lane,
	\emph{Categories for the Working Mathematician}, 2nd ed.,
	Graduate Texts in Mathematics \textbf{5}
	(Springer, 1998).
	\bibitem{Rudin1987}
	W.~Rudin,
	\emph{Real and Complex Analysis}, 3rd ed.
	(McGraw--Hill, 1987).
	\bibitem{Lee2013}
	J.~M.~Lee,
	\emph{Introduction to Smooth Manifolds}, 2nd ed.,
	Graduate Texts in Mathematics \textbf{218}
	(Springer, 2013).
	\bibitem{KobayashiNomizu1963}
	S.~Kobayashi and K.~Nomizu,
	\emph{Foundations of Differential Geometry}, Vol.~I
	(Interscience, 1963).
	\bibitem{Wald1984}
	R.~M.~Wald,
	\emph{General Relativity}
	(University of Chicago Press, 1984).
	\bibitem{VonNeumann1955}
	J.~von~Neumann,
	\emph{Mathematical Foundations of Quantum Mechanics}
	(Princeton University Press, 1955).
	\bibitem{Kato1995}
	T.~Kato,
	\emph{Perturbation Theory for Linear Operators},
	Classics in Mathematics
	(Springer, 1995).
	\bibitem{Hall1959}
	M.~Hall, Jr.,
	\emph{The Theory of Groups}
	(Macmillan, 1959).
	\bibitem{Gleason1952}
	A.~M.~Gleason,
	``Groups without small subgroups,''
	Ann.\ of Math.\ \textbf{56} (1952), 193--212.
	\bibitem{Yamabe1953}
	H.~Yamabe,
	``A generalization of a theorem of Gleason,''
	Ann.\ of Math.\ \textbf{58} (1953), 351--365.
	\bibitem{MontgomeryZippin1955}
	D.~Montgomery and L.~Zippin,
	\emph{Topological Transformation Groups}
	(Interscience, 1955).
	\bibitem{vanDantzig1936}
	D.~van Dantzig,
	``Zur topologischen Algebra.\ III.\ Brouwersche und Cantorsche
	Gruppen,''
	Compositio Math.\ \textbf{3} (1936), 408--426.
	\bibitem{Tao2014}
	T.~Tao,
	\emph{Hilbert's Fifth Problem and Related Topics},
	Graduate Studies in Mathematics \textbf{153}
	(American Mathematical Society, 2014).
	\bibitem{Malcev1949}
	A.~I.~Mal'cev,
	``On a class of homogeneous spaces,''
	Izv.\ Akad.\ Nauk SSSR Ser.\ Mat.\ \textbf{13} (1949), 9--32.
	\bibitem{Magnus1935}
	W.~Magnus,
	``Beziehungen zwischen Gruppen und Idealen in einem speziellen
	Ring,''
	Math.\ Ann.\ \textbf{111} (1935), 259--280.
	\bibitem{Witt1937}
	E.~Witt,
	``Treue Darstellung Liescher Ringe,''
	J.\ Reine Angew.\ Math.\ \textbf{177} (1937), 152--160.
	\bibitem{Labute1970}
	J.~P.~Labute,
	``On the descending central series of groups with a single defining
	relation,''
	J.\ Algebra \textbf{14} (1970), 16--23.
	\bibitem{DuchampKrob1992}
	G.~Duchamp and D.~Krob,
	``The lower central series of the free partially commutative
	group,''
	Semigroup Forum \textbf{45} (1992), 385--394.
	\bibitem{Holder1901}
	O.~H\"older,
	``Die Axiome der Quantit\"at und die Lehre vom Mass,''
	Ber.\ Verh.\ S\"achs.\ Ges.\ Wiss.\ Leipzig Math.-Phys.\ Cl.\
	\textbf{53} (1901), 1--64.
	\bibitem{Peetre1959}
	J.~Peetre,
	``Une caract\'erisation abstraite des op\'erateurs
	diff\'erentiels,''
	Math.\ Scand.\ \textbf{7} (1959), 211--218;
	rectification, Math.\ Scand.\ \textbf{8} (1960), 116--120.
	\bibitem{EpsteinThurston1979}
	D.~B.~A.~Epstein and W.~P.~Thurston,
	``Transformation groups and natural bundles,''
	Proc.\ London Math.\ Soc.\ (3) \textbf{38} (1979), 219--236.
	\bibitem{Lovelock1971}
	D.~Lovelock,
	``The Einstein tensor and its generalizations,''
	J.\ Math.\ Phys.\ \textbf{12} (1971), 498--501.
	\bibitem{CheegerMullerSchrader1984}
	J.~Cheeger, W.~M\"uller, and R.~Schrader,
	``On the curvature of piecewise flat spaces,''
	Comm.\ Math.\ Phys.\ \textbf{92} (1984), 405--454.
	\bibitem{HojmanKucharTeitelboim1976}
	S.~A.~Hojman, K.~Kucha\v{r}, and C.~Teitelboim,
	``Geometrodynamics regained,''
	Ann.\ Phys.\ (N.Y.) \textbf{96} (1976), 88--135.
	\bibitem{Nakahara2003}
	M.~Nakahara,
	\emph{Geometry, Topology and Physics}, 2nd ed.
	(Institute of Physics, 2003).
	\bibitem{Wilson1974}
	K.~G.~Wilson,
	``Confinement of quarks,''
	Phys.\ Rev.\ D \textbf{10} (1974), 2445--2459.
	DOI: \href{https://doi.org/10.1103/PhysRevD.10.2445}{10.1103/PhysRevD.10.2445}.
	\bibitem{Northrop1963}
	T.~G.~Northrop,
	\emph{The Adiabatic Motion of Charged Particles}
	(Interscience, 1963).
	\bibitem{RepovsScepin1997}
	D.~Repov\v{s} and E.~V.~\v{S}\v{c}epin,
	``A proof of the Hilbert--Smith conjecture for actions by Lipschitz
	maps,''
	Math.\ Ann.\ \textbf{308} (1997), 361--364.
	\bibitem{Pardon2013}
	J.~Pardon,
	``The Hilbert--Smith conjecture for three-manifolds,''
	J.\ Amer.\ Math.\ Soc.\ \textbf{26} (2013), 879--899.

	\bibitem{SingerWermer1955}
	I.~M.~Singer and J.~Wermer,
	``Derivations on commutative normed algebras,''
	Math.\ Ann.\ \textbf{129} (1955), 260--264.
	DOI: \href{https://doi.org/10.1007/BF01362370}{10.1007/BF01362370}.

	\bibitem{Connes2013}
	A.~Connes,
	``On the spectral characterization of manifolds,''
	J.\ Noncommut.\ Geom.\ \textbf{7} (2013), no.~1, 1--82.
	DOI: \href{https://doi.org/10.4171/JNCG/108}{10.4171/JNCG/108}.
\end{thebibliography}
\end{document}
