% !TeX root = closed-systems-from-comparison-completeness.tex
% ============================================================
\chapter*{List of Principal Results}
\phantomsection
\addcontentsline{toc}{chapter}{List of Principal Results}
\label{ch:principal-results-list}
% ============================================================
This list records the main formal waypoints in dependency order.  It is
not a complete list of theorem-like statements; local definitions,
lemmas, propositions, corollaries, and auxiliary theorems remain in the
chapters where they are proved.  The detailed back-matter result index
records the principal results again by topic.

\newcommand{\resultloc}[1]{p.~\pageref{#1}}
\newcommand{\resultitem}[3]{%
  \item[#1] #3 See \cref{#2}, \resultloc{#2}.%
}

\section*{Comparison and Quotient Backbone}
\begin{description}[leftmargin=4.8cm,style=nextline,itemsep=0.25em,before=\raggedright]
\resultitem{Structural rigidity}{thm:intro-structural-rigidity}
{The closed relational program is stated as a dependency theorem from
comparison data through quotient, transport, realization, and
state-side reconstruction.}
\resultitem{Quantifier-reversal boundary}{thm:boundary}
{The finite-to-global obstruction for admissible refinement towers is
isolated in Boolean form.}
\resultitem{Comparison-world classification}{thm:classification}
{Comparison worlds are classified by their intrinsic profile data before
additional geometric or field-theoretic structure is introduced.}
\resultitem{Rectangular completeness}{thm:rectangular-completeness-intrinsicality}
{Rectangular completeness is derived from intrinsicality and local
distinguishability rather than imposed as a background product.}
\resultitem{Canonical descent}{thm:descent}
{Coherent reports descend through the intrinsic quotient.}
\resultitem{Two-locus classification}{thm:two-locus-classification}
{Any quotient-semantic extension decomposes into representative choice,
transport, or both.}
\resultitem{Lift classification}{thm:lift-classification-main}
{Representative lifts are classified by representative data together
with compatible transport cocycles.}
\end{description}

\section*{Transport, Curvature, and Geometry}
\begin{description}[leftmargin=4.8cm,style=nextline,itemsep=0.25em,before=\raggedright]
\resultitem{Triangle regime}{thm:triangle}
{The first irreducible multi-point obstruction occurs at the triangle
level.}
\resultitem{Descent obstruction}{thm:DO-classification}
{Loop holonomy classifies the obstruction left by representative lifts.}
\resultitem{Triangle--loop bridge}{thm:bridge-main}
{The finite triangle obstruction is identified with the global loop
defect at the relevant transport boundary.}
\resultitem{Intrinsic curvature}{thm:interface-curvature}
{The first stable quadratic transport carrier is identified with realized
curvature under the stated smooth interface.}
\resultitem{Hilbert reconstruction}{thm:hilbert-classification}
{The quadratic carrier determines the Hilbert-side structure in the
specified realization regime.}
\resultitem{Quadratic scalar}{thm:quadratic-scalar}
{A positive quadratic scalar channel is isolated on the stabilized
carrier.}
\resultitem{Einstein boundary}{thm:einstein-main}
{Macroscopic compatibility yields the Einstein-form boundary law under
the stated realization hypotheses.}
\resultitem{Connection-first geometry}{thm:first-order-classification}
{First-order geometry is recovered as the intrinsic first-order
transport residue.}
\end{description}

\section*{Phase, Matter, Numerical, and State-Side Results}
\begin{description}[leftmargin=4.8cm,style=nextline,itemsep=0.25em,before=\raggedright]
\resultitem{Phase reduction}{thm:em-phase-reduction}
{The first visible phase sector reduces to \(U(1)\).}
\resultitem{Charge lattice}{thm:em-charge-lattice}
{The fundamental closed-system charge lattice is fixed at the primitive
phase-sector level.}
\resultitem{Denominator-\(3\) uniqueness}{thm:em-denominator3-uniqueness}
{Composite fractional charge descends through a minimal denominator-\(3\)
quotient.}
\resultitem{Matter-sector scalar}{thm:spectral-scalar-identification}
{Primitive matter sectors inherit a scalar invariant from the quadratic
channel.}
\resultitem{Mass-coordinate bridge}{thm:numbers-mass-calibration}
{The quartic coefficient-energy readout is identified with the intrinsic
mass-coordinate bridge.}
\resultitem{Maximal intrinsic closure}{thm:final}
{The scalar-channel closure statement is consolidated at the end of the
numerical and shell-closure stack.}
\resultitem{Raw inexhaustibility}{thm:dyn-raw-inexhaustible}
{The raw comparison limit contains more information than any selected
realized smooth reading.}
\resultitem{Conditional completion ledger}{thm:cond-closure-normal-form}
{The remaining global quantitative dependencies are put into an explicit
normal form.}
\resultitem{Transport appendix invariant}{thm:fusion-main-statement}
{The appendix records the nonperturbative confinement-side invariant
package induced by the closed transport stack.}
\end{description}

\let\resultloc\relax
\let\resultitem\relax
