% !TeX root = closed-systems-from-comparison-completeness.tex
% ============================================================
% Publication, citation, and rights record
% ============================================================
\clearpage
\thispagestyle{empty}
\begingroup
\small
\raggedright
\vspace*{0.65in}

\noindent{\Large\bfseries Publication Record}

\vspace{0.22in}

\noindent
\textbf{Title.}
\textit{Closed Systems from Comparison Completeness: A Theory of Closed
Relational Systems}

\vspace{0.10in}

\noindent
\textbf{Author.}
Chast K.~Wolfe

\vspace{0.10in}

\noindent
\textbf{Affiliation.}
Closure Research Initiative, United States

\vspace{0.10in}

\noindent
\textbf{Correspondence.}
\href{mailto:chast.wolfe@closureresearchinitiative.org}
{\texttt{chast.wolfe@closureresearchinitiative.org}}

\vspace{0.10in}

\noindent
\textbf{Canonical site.}
\href{https://closureresearchinitiative.org/csm/}
{\texttt{https://closureresearchinitiative.org/csm/}}

\vspace{0.10in}

\noindent
\textbf{Version.}
Version 11, June 14, 2026.

\vspace{0.10in}

\noindent
\textbf{DOI status.}
DOI pending for the current version. Earlier DOI records are assigned to
the versions listed in the public version history.

\vspace{0.10in}

\noindent
\textbf{Preferred citation.}
Chast K.~Wolfe,
\textit{Closed Systems from Comparison Completeness: A Theory of Closed
Relational Systems}, Version 11, Closure Research Initiative, 2026.
\href{https://closureresearchinitiative.org/csm/}
{\texttt{https://closureresearchinitiative.org/csm/}}.

\vspace{0.10in}

\noindent
\textbf{Version archive.}
The current release, archived PDFs, source packages, and version-specific
DOI records are maintained at
\href{https://closureresearchinitiative.org/csm/}
{\texttt{https://closureresearchinitiative.org/csm/}}.
Archived versions are fixed records of the versions to which they are
assigned.

\vspace{0.10in}

\noindent
\textbf{Rights.}
Copyright \(\copyright\) 2026 Chast K.~Wolfe. All rights reserved.
This manuscript is publicly available for reading, citation, criticism,
and review. No open-source license, public-domain dedication, or permission
to redistribute derivative works is granted by public availability. Uses
beyond ordinary academic citation, quotation, review, or fair-use/fair-dealing
contexts require prior written permission.

\vspace{0.10in}

\noindent
\textbf{Source record.}
Public source material, citation metadata, and rights information are
maintained at
\href{https://closureresearchinitiative.org/sources/}
{\texttt{https://closureresearchinitiative.org/sources/}}
and
\href{https://closureresearchinitiative.org/license/}
{\texttt{https://closureresearchinitiative.org/license/}}.

\vfill
\noindent
\textit{This record is bibliographic only; it does not state or modify any
mathematical result.}

\endgroup
\clearpage
