% !TeX root = closed-systems-from-comparison-completeness.tex
% ============================================================
% Declarations and source availability
% ============================================================
\clearpage
\thispagestyle{empty}
\begingroup
\small
\raggedright
\vspace*{0.65in}

\noindent{\Large\bfseries Declarations and Source Availability}

\vspace{0.22in}

\noindent
\textbf{Author.}
Chast K.~Wolfe.

\vspace{0.10in}

\noindent
\textbf{Affiliation and correspondence.}
Closure Research Initiative, United States;
\href{mailto:chast.wolfe@closureresearchinitiative.org}
{\texttt{chast.wolfe@closureresearchinitiative.org}}.

\vspace{0.10in}

\noindent
\textbf{Funding.}
No external grant or institutional funding is associated with this
version of the manuscript.

\vspace{0.10in}

\noindent
\textbf{Competing interests.}
No competing interests are declared.

\vspace{0.10in}

\noindent
\textbf{Data availability.}
This manuscript does not introduce an empirical dataset.  The data
objects studied in the text are mathematical structures defined in the
manuscript.

\vspace{0.10in}

\noindent
\textbf{Code availability.}
No numerical code is required to verify the definitions, lemmas, or
theorems stated in the manuscript.  Public website and manuscript source
packages are maintained through the Closure Research Initiative public
record.

\vspace{0.10in}

\noindent
\textbf{Source availability.}
The current PDF, versioned PDFs, and source bundles are maintained at
\href{https://closureresearchinitiative.org/csm/}
{\texttt{https://closureresearchinitiative.org/csm/}}.  Citation metadata
and external-source records are maintained at
\href{https://closureresearchinitiative.org/sources/}
{\texttt{https://closureresearchinitiative.org/sources/}}.

\vspace{0.10in}

\noindent
\textbf{Review status.}
This is a public versioned research manuscript.  Unless a later record
states otherwise, it should not be described as peer reviewed.

\vfill
\noindent
\textit{This declarations page is administrative.  It does not alter the
definitions, hypotheses, or results in the body of the manuscript.}

\endgroup
\clearpage
